---
title: "HACKATHON 2021"
output: 
  distill::distill_article:
    toc: true     # adding table of contents
    toc_depth: 2
    toc_float: true
    theme: theme.css
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web. 

Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.

```{css, echo=FALSE}



```
<!-- 
highest header (h1) styling (Header indicated by #)
subheader (h2) styling (Header indicated by ##)
body = background
vw is for responsive font size
-->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bslib)
```


# Themes
Themes to choosefrom: <br>
flatly, darkly, readable, spacelab, united, cosmo, lumen, paper, sandstone, simplex, and yeti.

```{r cars}

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Another header 1 that is really really loooooong
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>

# Another header 1
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
# Another header 1
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>

# Another header 1
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>

# Another header 1
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>

## Data Overview {.tabset}

<!-- tabset-pills is the background color of the button
.tabset-fade indicates whether the button should fade when clicking on the button
.tabset is the command to create a button/tab
-->
### Button1

Although the number of unique job ads is fairly modest, there is some very rich data to explore. We have over a year's worth of data, ranging between

```{r}

```


### Button2

Summary for each variable (only included R and Python here, as other languages (along with R and Python) are explored more deeply later.

```{r}

```


### Button3

Some functions are defined here for later use.

### Button4 with subtabs {.tabset}

#### Tab Nr 1{.tabset}
(tab1 content)

##### Subtab
(subtab content)

#### Functioning Tab
Summary of XY
<!-- Important to note: Buttons need to be one level below the title
-->

```{r panelset, echo=FALSE}
library(devtools)
devtools::install_github("gadenbuie/xaringanExtra")
xaringanExtra::use_panelset()
```

::::: {.panelset}

::: {.panel}

## X {.panel-name}

Amet enim aptent molestie vulputate pharetra
vulputate primis et vivamus semper.

:::

::: {.panel}

## Y {.panel-name}

### Sub heading one

Sit etiam malesuada arcu fusce ullamcorper
interdum proin tincidunt curabitur felis?

:::

::: {.panel}

## Tab Three {.panel-name}

Adipiscing mauris egestas vitae pretium 
ad dignissim dictumst platea!

:::

:::::





