---
title: "Hacklab Developer Census 2020"
description: "Preliminary Draft - `r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_float: true # not sure why not working
    toc_depth: 1
    # toc_collapsed: true
    theme: css_theme/theme.css
---




```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE) # we do not want to display the code

# Environment:

library(xaringanExtra) # to be able to have tabs. To install: install 'devtools' package, then run: devtools::install_github("gadenbuie/xaringanExtra")
xaringanExtra::use_panelset() # allow the creation of tabs in Distill (see https://github.com/rstudio/distill/issues/11)

library(tidyverse) # includes dplyr, tidyr, ggplot2, rio, stringr, ...

library(leaflet) # for the interactive map
library(tmaptools) # for geocoding

library(wordcloud) # for the wordcloud at Q33

library(tidyr) # for Q30 and Q31

```

```{r}

# Data:

Qs <- rio::import("data/clean/clean_all_Qs.rds") %>%
  # reodering certain factors for the plots: (use barplot_ordered() for plotting)
  mutate(
    purch_influence_16 = fct_relevel(purch_influence_16,
                                     "I have little or no influence",
                                     "I have some influence",
                                     "I have a great deal of influence"
    ),    
    job_satisfaction_22 = fct_relevel(job_satisfaction_22,
                                      "Very dissatisfied",
                                      "Slightly dissatisfied",
                                      "Neither satisfied nor dissatisfied",
                                      "Slightly satisfied",
                                      "Very satisfied"),
    edu_importance_20 = fct_relevel(edu_importance_20,
                                    "Not at all important/not necessary",
                                    "Somewhat important",
                                    "Fairly important",
                                    "Very important",
                                    "Critically important"),
    company_size_23 = fct_relevel(company_size_23,
                                  "Over 500",
                                  "Over 100",
                                  "Below 100",
                                  "Below 20",
                                  "Below 10",
                                  "1"),
    overtime_work_25 = fct_relevel(overtime_work_25,
                                   "Often: 1-2 days per week",
                                   "Frequently: 3 or more days per week",
                                   "Sometimes: 1-2 days per month but less than weekly",
                                   "Occasionally: 1-2 days per quarter but less than monthly",
                                   "Rarely: 1-2 days per year or less",
                                   "Never"),
    monthly_salary_28 = fct_relevel(monthly_salary_28,
                                    "Greater than GHS 25,000",
                                    "GHS 20,000 - GHS 25,000",
                                    "GHS 15,000 - GHS 20,000",
                                    "GHS 10,000 - GHS 15,000",
                                    "GHS 8,000 - GHS 10,000",
                                    "GHS 6,000 - GHS 8,000",
                                    "GHS 5,000 - GHS 6,000",
                                    "GHS 4,000 - GHS 5,000",
                                    "GHS 3,500 - GHS 4,000",
                                    "GHS 3,000 - GHS 3,500",
                                    "GHS 2,500 - GHS 3,000",
                                    "GHS 2,000 - GHS 2,500",
                                    "GHS 1,500 - GHS 2,000",
                                    "Less than GHS 1,500"),
    prim_study_19 = recode(prim_study_19, 
                           "Computer_Science_Engineering" = "Computer Science/Engineering",
                           "Information_Technology" = "Information Technology",
                           "Other_Engineering" = "Other Engineering",
                           "Health_Science" = "Health Science",
                           "Social_Science" = "Social Science",
                           "Web_Development_Web_Design" = "Web-Development/Web-Design",
                           "General_Science" = "General Science",
                           "Electrical_Engineering" = "Electrical Engineering",
                           "Visual_Arts" = "Visual Arts",
                           "Agricultural_Science" = "Agricultural Science",
                           "Graphic_Design" = "Graphic Design",
                           "General_Arts" = "General Arts",
                           "Mathematics_Statistics" = "Mathematics/Statistics"),
    prim_study_19 = fct_relevel(prim_study_19,
                                "Education",
                                "Social Science",
                                "Graphic Design",
                                "Architecture",
                                "Visual Arts",
                                "Business",
                                "General Arts",
                                "General Science",
                                "Health Science",
                                "Chemistry",
                                "Other Engineering",
                                "Physics",
                                "Agricultural Science",
                                "Web-Development/Web-Design",
                                "Mathematics/Statistics",
                                "Electrical Engineering",
                                "Information Technology",
                                "Computer Science/Engineering"),   
    highest_edu_18 = recode(highest_edu_18, 
                            "Secondary_High_School" = "Secondary High School",
                            "Basic_Education" = "Basic Education",
                            "Higher_National_Diploma" = "Higher National Diploma",
                            "Professional_Diploma" = "Professional Diploma",
                            "Teacher_Diploma" = "Teacher Diploma",
                            "College" = "Bachelor"),
    highest_edu_18 = fct_relevel(highest_edu_18,
                                 "Basic Education",
                                 "Secondary High School",
                                 "Teacher Diploma",
                                 "Professional Diploma",
                                 "Higher National Diploma",
                                 "Bachelor",
                                 "Master")
  )

skills <- rio::import("data/clean/skills_final.csv") %>% 
  select(-V1) %>%  # artifact due to a save as .csv to remove (at least on a Mac?)
  mutate_all(na_if,"") # the column 'level' contains blanks ("") instead of NAs -> transform them to NAs.
# we add characteristics of the respondents to the skills, 
# so that we can differentiate their popularity by occupation or gender:
skills_with_characteristics <- skills %>% left_join(select(Qs, ID, profession_1, gender_35), by = c("id" = "ID"))
 
number_of_respondents = dim(Qs)[1] # number of respondents (272) used to compute percentages.
number_of_prof_dev = dim(Qs %>% filter(profession_1 == "I am a developer by profession"))[1]
number_of_students = dim(Qs %>% filter(profession_1 == "I am a student who is learning to code"))[1]


```

```{r}

# Themes and Functions:

HL_colors = c("#610b70","#88b101","#eb1c96","#e98403","#454545")
HL_font = "sans" # !!! Should change and use Gotham !!!
main_fontsize = 7 

barplot_theme <- function(){
  # the definition of a custom ggplot theme for our barplots.
  theme_minimal() + 
    theme(
      plot.title = element_text(family = HL_font, size = main_fontsize + 2, color = "black"),
      # plot.subtitle = element_text(family = HL_font, size = main_fontsize),
      axis.title = element_text(family = HL_font, size = main_fontsize, color = "black"),
      axis.text = element_text(family = HL_font, size = main_fontsize, color = "black"),
      # plot.caption = element_text(hjust = 0, family = my_font, 
      #                             size = main_fontsize, face= "italic"), #Default is hjust=1 / 0
      plot.title.position = "plot",  # left-align title
      # plot.caption.position =  "plot", 
      panel.grid.minor.x = element_blank(),  # remove useless minor grid
      panel.grid.major.y = element_blank(),  # remove useless major grid in bar charts
      panel.grid.major.x = element_blank(),  # remove useless major grid in bar charts
      axis.title.x = element_blank(), # we do not want an y axis for the barplot, we have labels already
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
      # plot.margin = margin(0.1, 0.1, 0.1, 0.1, unit = "cm")
    )
}

basic_barplot <- function(my_df, a_factor, its_values, labels, title){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = fct_reorder({{a_factor}}, {{its_values}}), 
                             y = {{its_values}})) + 
    geom_col(fill = HL_colors[1]) + 
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.06*max({{its_values}}))),
      size = 2
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}

#Function for barplot that does not override order
barplot_ordered <- function(my_df, a_factor, its_values, labels, title){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = {{a_factor}},
                             y = {{its_values}})) + 
    geom_col(fill = HL_colors[1]) + 
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.06*max({{its_values}}))),
      size = 2
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}

compute_perc <- function(my_df, a_factor){
  # a small function to count a factor and compute the percentages
  # also return the percentages as char for labelling.
  # USE THIS FUNCTION IF THE SUM OF THE ANSWERS IS 272! (including NAs) -> not working if several possible choices.
  my_df %>%
    count({{a_factor}}) %>%
    drop_na() %>% # always in percentage of the respondents to the question.
    mutate(
      perc = n/sum(n),
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% return()
  
}

compute_perc_popular_skills <- function(the_skills, skillset, number_of_resp){
  # USE THIS FUNCTION for the popular skills (last year + both)
  # NOTE: we use all respondents, not only those who responded. -> some respondents did not enter any skills. -> not optimal, but annoying to correct?
  the_skills %>%
    filter(tool %in% skillset) %>% # select only the languages from all the tools
    filter(level == "Worked with in PAST year" | level == "Both") %>%
    count(tool) %>%
    mutate(
      perc = n/number_of_resp, 
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% 
    arrange(-perc) %>%  
    return()
}

compute_perc_future_skills <- function(the_skills, skillset, number_of_resp){
  # USE THIS FUNCTION for the future, nexy year, skills (next year + both)
  # NOTE: we use all respondents, not only those who responded. -> some respondents did not enter any skills. -> not optimal, but annoying to correct?
  the_skills %>%
    filter(tool %in% skillset) %>% # select only the languages from all the tools
    filter(level == "Want to work with NEXT year" | level == "Both") %>%
    count(tool) %>%
    mutate(
      perc = n/number_of_resp, 
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% 
    arrange(-perc) %>%  
    return()
}


```


# Basic Information

## What describes you best?

The two biggest subgroups among the respondents are professional developers as well as students. Additionally, there are respondents coding as a part of their work, coding as a hobby, as well as former developers. 

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, profession_1), 
              a_factor = profession_1, its_values = perc, labels = perc_label, 
              title = "")
```

## Do you code as a hobby? 

Most of the respondents code as a hobby. Interestingly, professional developers seem to code for a hobby less often than students which may be related to their reduced time availability. 

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::

## Employment

```{r}

employment_data <- Qs %>%
  filter(employment_3 != "I prefer not to say") %>%
  mutate(
    employment_3 = factor(employment_3, 
                          levels(employment_3)[c(4,2,3,6,5,1)])
  )
```

Most of the respondents are employed full-time. Furthermore, there is a great share of students. Sadly, there is also a significant share of unemployed developers looking for work. 

When it comes to gender differences, there is a greater share of students among the female respondents. In addition, women seem to be self-employed less often when compared with men. 

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(employment_data, employment_3), 
              a_factor = employment_3, its_values = perc, labels = perc_label,
              title = "")

```


:::

::: {.panel}

### Men only {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(employment_data %>% filter(gender_35 == "Man"), employment_3), 
              a_factor = employment_3, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Women only {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(employment_data %>% filter(gender_35 == "Woman"), employment_3), 
              a_factor = employment_3, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::



## Where do the developers come from? 

### Region

Most respondents come from the greater Accra region. However, the extent of concentration in Accra seems to be larger for professional developers than for students. 

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}
region_data <- Qs %>%
  filter(region_4 != "Not in Ghana")

region_data$region_4 <- recode(region_data$region_4,
         "Eastern Region, Ghana" = "Eastern Region", 
         "Northern Region, Ghana" = "Northern Region", 
         "Central Region, Ghana" = "Central Region", 
         "Western Region, Ghana"= "Western Region"
  )

basic_barplot(my_df = compute_perc(region_data, region_4), 
              a_factor = region_4, its_values = perc, labels = perc_label,
              title = "")

```


:::

::: {.panel}

### Professional Developers {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(region_data %>% filter(profession_1 == "I am a developer by profession"), region_4), 
              a_factor = region_4, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(region_data %>% filter(profession_1 == "I am a student who is learning to code"), region_4), 
              a_factor = region_4, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::




### City

Zoom in to get a more detailed picture. 

Hover over the individual markers to see the professional status of the respondents. 

```{r, fig.align= 'center', out.width= '100%', fig.height=6.5}

# the geocoding was done separately in data_cleaning/Geocoding.R
# (to not lose time everytime we knitr)
# we load the result:
data_city_geo <- rio::import("data/clean/data_city_geo.rds") 

#Map for the cities
violet_icon <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/pointhi/leaflet-color-markers/master/img/marker-icon-violet.png", 
  iconWidth = 24,
  iconHeight = 32) 

leaflet(data = data_city_geo) %>%
  addTiles() %>% 
  addMarkers(~ geocode$lon, ~ geocode$lat, 
             clusterOptions = markerClusterOptions(), 
             icon = violet_icon, label = ~profession_1)
```

## How old are the respondents? 

Most respondents are between 20 and 30. 

While the students are very young in most cases, the professional developers display a little more variance in their ages. However, we have a very young sample at hand.

```{r}
Qs$age_range_6 <- factor(Qs$age_range_6, levels(Qs$age_range_6)[c(6:1)])

```

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, age_range_6), 
              a_factor = age_range_6, its_values = perc, labels = perc_label, 
              title = "")

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}


barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), age_range_6), 
              a_factor = age_range_6, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), age_range_6), 
              a_factor = age_range_6, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::







# Technology and Tech Culture

## How frequently do you learn a new language or framework?

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, new_tool_7), 
              a_factor = new_tool_7, its_values = perc, labels = perc_label, 
              title = "") + scale_x_discrete(limits=rev)

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), new_tool_7), 
              a_factor = new_tool_7, its_values = perc, labels = perc_label, 
              title = "") + scale_x_discrete(limits=rev)

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), new_tool_7), 
              a_factor = new_tool_7, its_values = perc, labels = perc_label, 
              title = "") + scale_x_discrete(limits=rev)

```

:::

:::::



## Which programming, scripting and markup languages have you done extensive development work in over the past year?

_HTML/CSS_ is the most used language, followed by _JavaScript_, _Python_ and _SQL_. Professional developers are more likely to use _JavaScript_ and _SQL_. _C++_ and _C_ are still widely used to teach coding to students.    

<aside>
  note: includes "both" and "this year".
</aside>  

```{r}

languages = c(
  "C",			
  "C#",			
  "C++",		
  "HTML/CSS",			
  "Java",			
  "JavaScript",			
  'Kotlin',			
  "PHP",			
  "Python",			
  "Ruby",			
  "SQL",			
  "Swift",			
  "TypeScript"
)


```
::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics, 
                                                  skillset = languages,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = languages,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = languages,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## Which platforms have you done extensive development work in over the past year?

A comment here...  

```{r}

platforms = c(
"Android",			
"Arduino",			
"AWS",			
"Docker",		
"Google Cloud Platform",			
"Heroku",			
"IBM Cloud or Watson",			
"iOS",			
"Linux",			
"MacOS",		
"Microsoft Azure",		
"Raspberry Pi", 	
"Windows",			
"WordPress"
)

```
::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics, 
                                                  skillset = platforms,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = platforms,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = platforms,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## Which web frameworks have you done extensive development work in over the past year?

_React.js_ is the most popular web framework.   

```{r}

web_frameworks = c(
"Angular",			
"ASP.NET",		
"Django",			
"Express",		
"jQuery",			
"Laravel",
"React.js",
"Ruby on Rails",
"Vue.js",
"Spring"
)

```
::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics, 
                                                  skillset = web_frameworks,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = web_frameworks,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = web_frameworks,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## Which other frameworks, libraries and tools have you done extensive development work in over the past year?

A comment here...  

```{r}

other_frameworks = c(
".NET",			
".NET Core",			
"Flutter",			
"Node.js",			
"Puppet",			
"React Native",			
"Deno.js"
)

```
::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics, 
                                                  skillset = other_frameworks,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = other_frameworks,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = other_frameworks,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## Which collaborative tools have you used extensively over the past year?

_GitHub_ is by far the most popular version control tool. _Slack_ is the most used business communication platform.  

```{r}

collaborative_tools = c(
"GitHub",			
"GitLab",			
"Facebook Workplace",			
"Slack",			
"Microsoft Teams",			
"Microsoft Azure",			
"Trello",			
"Google Suite (Docs, Meet, etc)",			
"Stack Overflow for Teams",			
"Jira"
)

```
::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics, 
                                                  skillset = collaborative_tools,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = collaborative_tools,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = collaborative_tools,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::



&nbsp;  

&nbsp;   

== __Not sure yet how to combine this year and next year, tbd. For now they are separated.__ ==

&nbsp;    

&nbsp;  

## Which programming, scripting and markup languages do you want to work in over the next year?

A comment here... 

<aside>
  note: includes "both" and "next year" (for the 5 following plots)
</aside>  


::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics, 
                                                  skillset = languages,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = languages,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = languages,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## Which platforms do you want to work with over the next year?

A comment here...  

::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics, 
                                                  skillset = platforms,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = platforms,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = platforms,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## Which web frameworks do you want to work with over the next year?

_React.js_ is the most popular web framework.   

```{r}

web_frameworks = c(
"Angular",			
"ASP.NET",		
"Django",			
"Express",		
"jQuery",			
"Laravel",
"React.js",
"Ruby on Rails",
"Vue.js",
"Spring"
)

```
::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics, 
                                                  skillset = web_frameworks,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = web_frameworks,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = web_frameworks,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## Which other frameworks, libraries and tools do you want to work with over the next year?

A comment here...  

```{r}

other_frameworks = c(
".NET",			
".NET Core",			
"Flutter",			
"Node.js",			
"Puppet",			
"React Native",			
"Deno.js"
)

```
::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics, 
                                                  skillset = other_frameworks,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = other_frameworks,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = other_frameworks,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## Which collaborative tools do you want to work with over the next year?

_GitHub_ is by far the most popular version control tool. _Slack_ is the most used business communication platform.  

```{r}

collaborative_tools = c(
"GitHub",			
"GitLab",			
"Facebook Workplace",			
"Slack",			
"Microsoft Teams",			
"Microsoft Azure",			
"Trello",			
"Google Suite (Docs, Meet, etc)",			
"Stack Overflow for Teams",			
"Jira"
)

```
::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics, 
                                                  skillset = collaborative_tools,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = collaborative_tools,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = collaborative_tools,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


&nbsp;  

&nbsp;  

&nbsp;  

&nbsp;  


## What is the primary operating system in which you work?

Windows in the most used operating system. Professional developers use MacOS and Linux more often than other respondents. Students are overwhelmingly on Windows.  

::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## In which operating system would you rather work?

MacOS, Linux and Windows are practically tied as preferred systems. Professional developers have a preference for MacOS, then Linux. Students prefer Windows.  

::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::



## What level of influence do you, personally, have over new technology purchases at your organization?

Unsurprisingly, professional developers are more likely to have influence over new technology purchases.  

::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, purch_influence_16), 
              a_factor = purch_influence_16, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% 
                                     filter(profession_1 == "I am a developer by profession"),
                                   purch_influence_16), 
              a_factor = purch_influence_16, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::

## Q17. When buying a new tool or software, how do you discover and research available solutions? Select all that apply.

Most of the respondents declared relying on free trials and colleagues.  

```{r, fig.align='center', out.width = '100%', fig.height=3}

# multiple choice question: we separate the different answers (separate_rows),
#   and compute the % according to the total number of respondents
perc_solution_research_17 <- Qs %>% 
  select(ID, profession_1, solution_research_17) %>% 
  separate_rows(solution_research_17, sep = ";") %>% 
  # drop_na(any_of("solution_research_17")) %>% 
  filter(solution_research_17 != "") %>% 
  mutate(
    solution_research_17 = ifelse(solution_research_17 == "YouTube Videos", "YouTube", solution_research_17), # group those
    solution_research_17 = ifelse(solution_research_17 == "YouTube ", "YouTube", solution_research_17) # group those
  ) %>%
  count(solution_research_17) %>% 
  mutate(
      perc = n/number_of_respondents,  # % to total number of respondents -> some resp. selected several choices
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% 
    arrange(-perc) 

basic_barplot(my_df = perc_solution_research_17, 
              a_factor = solution_research_17, its_values = perc, labels = perc_label, 
              title = "")

  

```










# Education, Work and Career

## Q18. Which of the following best describes the highest level of formal education that you have completed?

Most of the survey participants have at least a secondary high-school degree whereas half of the students have a secondary high school degree and every third professional developer has a bachelor's degree

::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, highest_edu_18), 
              a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), highest_edu_18), 
              a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), highest_edu_18), 
              a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## Q19. What was your primary field of study?

Most participants study or studied Computer Science or Computer Engineering, followed by Information Technology and other Engineering fields as well as Mathematics/Statistics and Business.

::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, prim_study_19), 
              a_factor = prim_study_19, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), prim_study_19), 
              a_factor = prim_study_19, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), prim_study_19), 
              a_factor = prim_study_19, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## Q20. How important is a formal education, such as a university degree in computer science, to your career?

Half of the respondents find a formal education very or even critically important for their career.

::::: {.panelset}
::: {.panel}
### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, edu_importance_20), 
              a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), edu_importance_20), 
              a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), edu_importance_20), 
              a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


## Q21. From Q18, if you could go back and change your educational path (but end up in the same career), what would you change?

```{r, fig.align='center', out.width = '100%', fig.height=3}

compute_perc(Qs, change_edu_21 )

```

## Q22. How satisfied are you with your current job? (If you work multiple jobs, answer for one you spend the most hours on.)

A little bit than more of 50% of the respondents are slightly satisfied or satisfied with their current jobs.
Whereas 20% are slightly dissatisfied or dissatisfied.

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, job_satisfaction_22 ), 
              a_factor = job_satisfaction_22 , its_values = perc, labels = perc_label, 
              title = "Q22. How satisfied are you with your current job?")


```

## Relationship between job satisfaction and overtime work

Interestingly, people who work more than 3 times a week overtime are satisfied with their job. There seems no relationship that the more overtime the less job satisfaction. The tendency is rather the opposite. 

```{r, fig.align='center', out.width = '100%', fig.height=3}

# Heat Map: Relationship between job satisfaction and overtime work
runningcounts.df <- as.data.frame(table(Qs$job_satisfaction_22, Qs$overtime_work_25))

heatmap_Q22_Q25 <- runningcounts.df %>% rename(overtime_work_25 = Var2,
                                               job_satisfaction_22 = Var1) %>%
  mutate(overtime_work_25 = recode(overtime_work_25, 
                                                "Often: 1-2 days per week" = "1-2 days per week",
                                                "Sometimes: 1-2 days per month but less than weekly" = "1-2 days per month",
                                      "Rarely: 1-2 days per year or less" = "<1-2 days per year",
                                   "Occasionally: 1-2 days per quarter but less than monthly" = "1-2 days per quarter",
                                   "Frequently: 3 or more days per week" = ">3 days per week"))
                                      
levels(heatmap_Q22_Q25$overtime_work_25) <- gsub(" ", "\n", levels(heatmap_Q22_Q25$overtime_work_25))
levels(heatmap_Q22_Q25$job_satisfaction_22) <- gsub(" ", "\n", levels(heatmap_Q22_Q25$job_satisfaction_22))


                                      
ggplot(heatmap_Q22_Q25, aes(overtime_work_25, job_satisfaction_22)) +
  geom_tile(aes(fill = Freq), colour = "black") +
  scale_fill_gradient(low = "white", high = HL_colors[1])

# # Barplot: Relationship between job satisfaction and overtime work
# ggplot(runningcounts.df, aes(factor(Var1))) +
#   geom_bar(aes(y = Freq, alpha = rev(Var2)),
#            stat = "identity",
#            color = "#209f1b", fill = "#35665C") +
#   geom_text(aes(y = -5, label = Freq), color = "red") +
#   facet_grid(Var2 ~ Var1, switch = "both") +
#   labs(title = "frequency",
#        y = "Satsifaction",
#        x = "Overtime") +
#   theme_minimal() +
#   theme(strip.background = element_blank(),
#         axis.text.y = element_blank(),
#         axis.text.x = element_text(size = 12),
#         axis.ticks = element_blank(),
#         strip.text.y = element_text(angle = 0),
#         strip.text = element_text(size = 12),
#         legend.position = "none")

```

## Q23. Approximately how many people are employed by the company or organization you currently work for?

Half of the respondents work in a small company with less than 10 workers. Only 10% of the respondents work in a big company with more than 100 workers.

```{r, fig.align='center', out.width = '100%', fig.height=3}

#Removing NA

company_size_23_NA_remov <- Qs %>%
  filter(company_size_23 != "NA")

barplot_ordered(my_df = compute_perc(company_size_23_NA_remov, company_size_23 ), 
              a_factor = company_size_23 , its_values = perc, labels = perc_label, 
              title = "Q23. Approximately how many people are employed by the company or organization you currently work for?")

```

## Q24. On average, how many hours per week do you work? Please enter a whole number in the box.
Working hours are quite evenly scattered across the scale from 0-80 hours a week.

```{r, fig.align='center', out.width = '100%', fig.height=3}

ggplot(compute_perc(Qs,weekly_work_hours_24), aes(x=weekly_work_hours_24)) + geom_histogram(color = "white", fill= HL_colors[1], binwidth = 10) + labs(x="Weekly working hours", y = "Number of Respondents") +
    scale_x_continuous(breaks = seq(0, 120, by=10)) +
  theme_minimal() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
    # theme(
    #   plot.title = element_text(family = HL_font, size = main_fontsize + 2, color = "black"),
    #   # plot.subtitle = element_text(family = HL_font, size = main_fontsize),
    #   axis.title = element_text(family = HL_font, size = main_fontsize, color = "black"),
    #   axis.text = element_text(family = HL_font, size = main_fontsize, color = "black"),
    #   # plot.caption = element_text(hjust = 0, family = my_font, 
    #   #                             size = main_fontsize, face= "italic"), #Default is hjust=1 / 0
    #   plot.title.position = "plot",  # left-align title
    #   # plot.caption.position =  "plot", 
    #   panel.grid.minor.x = element_blank(),  # remove useless minor grid
    #   panel.grid.major.y = element_blank(),  # remove useless major grid in bar charts
    #   panel.grid.major.x = element_blank(),  # remove useless major grid in bar charts
    #   axis.text.x = element_blank(),
    #   axis.ticks.x = element_blank(),
    #   # plot.margin = margin(0.1, 0.1, 0.1, 0.1, unit = "cm")
    # )

```


## Q25. How often do you work overtime or beyond the formal time expectation of your job?

Around 50% of the respondents work overtime on 3 or more days in a week.

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, overtime_work_25 ), 
              a_factor = overtime_work_25 , its_values = perc, labels = perc_label, 
              title = "Q25. How often do you work overtime or beyond the formal time expectation of your job?")

```

## Q26. Do you think your company has a good on-boarding process? (By on-boarding, we mean the structured process of getting you settled in to your new role at a company)

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, company_onboarding_process_26 ), 
              a_factor = company_onboarding_process_26 , its_values = perc, labels = perc_label, 
              title = "26. Do you think your company has a good on-boarding process?")

```

## Q27. How could on-boarding at your company be improved?

```{r, fig.align='center', out.width = '100%', fig.height=3}

compute_perc(Qs, improve_onboarding_27 )

```

## Q28. How much is your monthly salary in Ghana Cedis?

50% of the respondents earn less than 2000 GHS per month. 

```{r}
monthly_salary_data <- Qs %>%
  filter(monthly_salary_28 != "NA")

```

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, monthly_salary_28 ), 
              a_factor = monthly_salary_28 , its_values = perc, labels = perc_label, 
              title = "")

```

:::

::: {.panel}

### Only Women {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(monthly_salary_data %>% filter(gender_35 == "Woman"), monthly_salary_28), 
              a_factor = monthly_salary_28, its_values = perc, labels = perc_label,
              title = "")

```
:::

::: {.panel}

### Only Men {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(monthly_salary_data %>% filter(gender_35 == "Man"), monthly_salary_28), 
              a_factor = monthly_salary_28, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::

## Q29. Which of the following best describes your current job-seeking status?

60% of the respondents are open for new job oppourtunities and one in three respondents is even actively looking for a job.

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, job_status_29 ), 
              a_factor = job_status_29 , its_values = perc, labels = perc_label, 
              title = "Q29. Which of the following best describes your current job-seeking status?")

```

## Q30. In general, what drives you to look for a new job? Select all that apply.
The main reason for developers is more income whereas for students it is gaining more experience and skills improvement.

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

drivers_for_new_jobs_30_split<-Qs%>% select(drivers_for_new_jobs_30, profession_1) %>%
  separate_rows(drivers_for_new_jobs_30, sep = ";")


drivers_for_new_jobs_30_split<-drivers_for_new_jobs_30_split[!(is.na(drivers_for_new_jobs_30_split$drivers_for_new_jobs_30) | drivers_for_new_jobs_30_split$drivers_for_new_jobs_30==""), ]
                                     
drivers_for_new_jobs_30_split <-drivers_for_new_jobs_30_split %>% 
  mutate( drivers_for_new_jobs_30 = recode(drivers_for_new_jobs_30, 
                                                "A job i love and how challenging it is" = "Other",
                                                "A Fulltime / Stable Employment " = "Other",
                                      "Building Credibility and Connections(Relationships). As well making my skills practical." = "Other",
                                                "Funds to seed stuffs" = "Other",
                                                "I am focussed on running my business." = "Other",
                                                "I have a self motivated personality" = "Other",
                                                "to be in an environment where am challenged beyond my current capabilities" = "Other"
))

basic_barplot(my_df = compute_perc(drivers_for_new_jobs_30_split, drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```
``

:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(drivers_for_new_jobs_30_split %>% filter(profession_1 == "I am a developer by profession"), drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```
:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(drivers_for_new_jobs_30_split %>% filter(profession_1 == "I am a student who is learning to code"), drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```
:::

::::


## Q31. When job searching, how do you learn more about a company? Select all that apply.

The most popular method to find out more about a company among the respondents is researching company media. There also seem no differences between developers and students.

```{r, fig.align='center', out.width = '100%', fig.height=3}

how_learn_about_company_31_split<-Qs%>% select(how_learn_about_company_31, profession_1) %>%
  separate_rows(how_learn_about_company_31, sep = ";")


how_learn_about_company_31_split<-how_learn_about_company_31_split[!(is.na(how_learn_about_company_31_split$how_learn_about_company_31) | how_learn_about_company_31_split$how_learn_about_company_31==""), ]


how_learn_about_company_31_split <-how_learn_about_company_31_split %>% 
  mutate(how_learn_about_company_31 = recode(how_learn_about_company_31, 
                                                "Company website" = "Read company media, such as employees blogs or company culture videos",
                                                "Google & their website" = "Read company media, such as employees blogs or company culture videos",
                                             "linkedin" = "Read company media, such as employees blogs or company culture videos",
                                             "Online (Linkedin" = "Read company media, such as employees blogs or company culture videos",
                                             "Websites " = "Read company media, such as employees blogs or company culture videos",
                                             "Personal assessment.. Occasionally going to the company to see how they cooperate with one another " = "Make personal contact to company",
                                             "Just going there frequently just access their interpersonal relationships with each other " = "Make personal contact with company"))
                                      
basic_barplot(my_df = compute_perc(how_learn_about_company_31_split, how_learn_about_company_31 ), 
              a_factor = how_learn_about_company_31 , its_values = perc, labels = perc_label, 
              title = "Q31. When job searching, how do you learn more about a company?")


```

## Q32. Are you a member of any other online/offline developer communities?

2 in 3 respondends are part of a developer community.

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, dev_community_member_32 ), 
              a_factor = dev_community_member_32 , its_values = perc, labels = perc_label, 
              title = "Q32. Are you a member of any other online/offline developer communities?")

```

## Q33. Developer communities you are a member of

The three largest communities to which respondents indicated membership were DevCongress (13.6%), Facebook Developer Circle (9.6%), and the Hacklab Foundation (7.4%).  

```{r, fig.align='center', out.width = '100%', fig.height=5.5}
# Warning! Check that all communities are in the plot. Depends on size (height) of figure and scale. Wordcloud excludes if names are too large (typically Facebook Developer Circle or DevCongress)
# I tried to pick a scale where the smallest were still legible.

# note: respondents can be members of several communities,
# we just count them all.
counted_communities <- Qs %>% select(ID, which_dev_community_33) %>%
  separate_rows(which_dev_community_33, sep = ";") %>% 
  drop_na() %>% 
  count(which_dev_community_33) %>% 
  arrange(-n) %>%
  mutate(perc_of_resp = n/number_of_respondents*100)

wordcloud(words = counted_communities$which_dev_community_33, freq = counted_communities$n, min.freq = 1,
          max.words = 122, random.order = TRUE, rot.per = 0, scale = c(2.2,0.5),
          colors = HL_colors)

```





# Demographics  

## Gender 

### What Gender do you identify as?

There is a  greater share of women among the students than among the professional developers. 

Ghana's developer community may become more representative in the upcoming years. 

```{r}

#Removing NA

gender_data <- Qs %>%
  filter(gender_35 != "NA")
```


::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(gender_data, gender_35), 
              a_factor = gender_35, its_values = perc, labels = perc_label, 
              title = "")

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}


barplot_ordered(my_df = compute_perc(gender_data %>% filter(profession_1 == "I am a developer by profession"), gender_35), 
              a_factor = gender_35, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(gender_data %>% filter(profession_1 == "I am a student who is learning to code"), gender_35), 
              a_factor = gender_35, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::


### Do you identify as transgender? 

```{r}
transgender_data <- Qs %>%
  filter(transgender_36 != "NA")
```


```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(transgender_data, transgender_36), 
              a_factor = transgender_36, its_values = perc, labels = perc_label,
              title = "")

```


## Sexual orientation

```{r}
sexual_orientation_data <- Qs %>%
    filter(sexual_orientation_37 != "NA" & sexual_orientation_37 != "Prefer not to say")

```

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(sexual_orientation_data, sexual_orientation_37), 
              a_factor = sexual_orientation_37, its_values = perc, labels = perc_label,
              title = "")

```

## Ethnicity 

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(ethnicity_38 != "NA" & ethnicity_38 != "I prefer not to answer"), ethnicity_38), 
              a_factor = ethnicity_38, its_values = perc, labels = perc_label,
              title = "")

```

## Do you have any dependents you care for? 

Among our respondents, women seem to have dependents they care for less often than men do. This may partly be due to the fact, that more women are still in their studies and that the women in our sample may be younger than the men. 

```{r}
dependents_data <- Qs %>%
  filter(dependents_41 != "I prefer not to say" & dependents_41 != "NA")
```

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(dependents_data, dependents_41), 
              a_factor = dependents_41, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Only Men {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(dependents_data %>% filter(gender_35 == "Man"), dependents_41), 
              a_factor = dependents_41, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Only Women {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(dependents_data %>% filter(gender_35 == "Woman"), dependents_41), 
              a_factor = dependents_41, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::




