---
title: "Hacklab Ghana Developer Census 2020"
subtitle: "Published by the Hacklab Foundation"
output: 
  html_document:  
    toc: true
    toc_float: true 
    toc_depth: 2
    toc_collapsed: true
    css: css_theme/theme.css # where the css theme is defined
    includes:
      after_body: css_theme/footer.html
---

```{css toc-content, echo = FALSE}
/* add some space before the TOC */
  /* for no apparent reason, this is working here but not in the .css... */
  #TOC {
  margin-top: 5rem;
}
```



```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warnings = FALSE,  # we do not want to display the code or warnings
                      dev = "svg")

# Environment:

library(xaringanExtra) # to be able to have tabs. To install: install 'devtools' package, then run: devtools::install_github("gadenbuie/xaringanExtra")
xaringanExtra::use_panelset() # allow the creation of tabs in Distill (see https://github.com/rstudio/distill/issues/11)

library(tidyverse) # includes dplyr, tidyr, ggplot2, rio, stringr, ...

library(leaflet) # for the interactive map
library(tmaptools) # for geocoding

library(wordcloud) # for the wordcloud at Q33

library(tidyr) # for Q30 and Q31

library(ggiraph) # for responsive bubble plots
library(plotly) # for interactive bubble plots

library(ggExtra) # for interactive Heatmaps
library(ggiraphExtra) # for interactive Heatmaps
library(moonBook) # for interactive Heatmaps
library(sjmisc) # for interactive Heatmaps

# library(ggchicklet) # to round edges of barplots. install.packages("ggchicklet", repos = "https://cinc.rud.is")


```

```{r}

# Data:
# reodering certain factors for the plots: (use barplot_ordered() for plotting)
Qs <- rio::import("data/clean/clean_all_Qs.rds") %>%
  mutate(
    purch_influence_16 = fct_relevel(purch_influence_16,
                                     "I have little or no influence",
                                     "I have some influence",
                                     "I have a great deal of influence"
    ),    
    job_satisfaction_22 = fct_relevel(job_satisfaction_22,
                                      "Very dissatisfied",
                                      "Slightly dissatisfied",
                                      "Neither satisfied nor dissatisfied",
                                      "Slightly satisfied",
                                      "Very satisfied"),
    edu_importance_20 = fct_relevel(edu_importance_20,
                                    "Not at all important/not necessary",
                                    "Somewhat important",
                                    "Fairly important",
                                    "Very important",
                                    "Critically important"),
    company_size_23 = fct_relevel(company_size_23,
                                  "Over 500",
                                  "Over 100",
                                  "Below 100",
                                  "Below 20",
                                  "Below 10",
                                  "1"),
    overtime_work_25 = fct_relevel(overtime_work_25,
                                   "Often: 1-2 days per week",
                                   "Frequently: 3 or more days per week",
                                   "Sometimes: 1-2 days per month but less than weekly",
                                   "Occasionally: 1-2 days per quarter but less than monthly",
                                   "Rarely: 1-2 days per year or less",
                                   "Never"),
    overtime_work_25 = recode(overtime_work_25,
                              "Often: 1-2 days per week" = "1-2 days per week",
                              "Sometimes: 1-2 days per month but less than weekly" = "1-2 days per month",
                              "Rarely: 1-2 days per year or less" = "<1-2 days per year",
                              "Occasionally: 1-2 days per quarter but less than monthly" = "1-2 days per quarter",
                              "Frequently: 3 or more days per week" = ">3 days per week"),
    monthly_salary_28 = fct_relevel(monthly_salary_28,
                                    "Greater than GHS 25,000",
                                    "GHS 20,000 - GHS 25,000",
                                    "GHS 15,000 - GHS 20,000",
                                    "GHS 10,000 - GHS 15,000",
                                    "GHS 8,000 - GHS 10,000",
                                    "GHS 6,000 - GHS 8,000",
                                    "GHS 5,000 - GHS 6,000",
                                    "GHS 4,000 - GHS 5,000",
                                    "GHS 3,500 - GHS 4,000",
                                    "GHS 3,000 - GHS 3,500",
                                    "GHS 2,500 - GHS 3,000",
                                    "GHS 2,000 - GHS 2,500",
                                    "GHS 1,500 - GHS 2,000",
                                    "Less than GHS 1,500"),
    prim_study_19 = recode(prim_study_19, 
                           "Computer_Science_Engineering" = "Computer Science/Engineering",
                           "Information_Technology" = "Information Technology",
                           "Other_Engineering" = "Other Engineering",
                           "Health_Science" = "Health Science",
                           "Social_Science" = "Social Science",
                           "Web_Development_Web_Design" = "Web-Development/Web-Design",
                           "General_Science" = "General Science",
                           "Electrical_Engineering" = "Electrical Engineering",
                           "Visual_Arts" = "Visual Arts",
                           "Agricultural_Science" = "Agricultural Science",
                           "Graphic_Design" = "Graphic Design",
                           "General_Arts" = "General Arts",
                           "Mathematics_Statistics" = "Mathematics/Statistics"),
    prim_study_19 = fct_relevel(prim_study_19,
                                "Education",
                                "Social Science",
                                "Graphic Design",
                                "Architecture",
                                "Visual Arts",
                                "Business",
                                "General Arts",
                                "General Science",
                                "Health Science",
                                "Chemistry",
                                "Other Engineering",
                                "Physics",
                                "Agricultural Science",
                                "Web-Development/Web-Design",
                                "Mathematics/Statistics",
                                "Electrical Engineering",
                                "Information Technology",
                                "Computer Science/Engineering"),   
    highest_edu_18 = recode(highest_edu_18, 
                            "Secondary_High_School" = "Secondary High School",
                            "Basic_Education" = "Basic Education",
                            "Higher_National_Diploma" = "Higher National Diploma",
                            "Professional_Diploma" = "Professional Diploma",
                            "Teacher_Diploma" = "Teacher Diploma",
                            "College" = "Bachelor"),
    highest_edu_18 = fct_relevel(highest_edu_18,
                                 "Basic Education",
                                 "Secondary High School",
                                 "Teacher Diploma",
                                 "Professional Diploma",
                                 "Higher National Diploma",
                                 "Bachelor",
                                 "Master"),
    profession_1 = recode(profession_1,
                          "I am a developer by profession" = "I am a developer by profession",
                          "I am not primarily a developer, but I write code sometimes as part of my work" = "I am not primarily a developer,\nbut sometimes write code at work",
                          "I am a student who is learning to code" = "I am a student who is learning to code",
                          "I code primarily as a hobby" = "I code primarily as a hobby",
                          "I used to be a developer by profession, but no longer am" = "I used to be a developer by profession,\nbut no longer am",
                          "None of these" = "None of these"),
    job_status_29 = as_factor(job_status_29),
    job_status_29 = recode(job_status_29,
                           "I'm not actively looking, but I am open to new opportunities" = "I'm not actively looking,\nbut I am open to new opportunities",
                           "I am actively looking for a job" = "I am actively looking for a job",
                           "I am not interested in new job oportunities" = "I am not interested\nin new job oportunities"),
    overtime_work_25 = recode(overtime_work_25,
                              "Often: 1-2 days per week" = "Often: 1-2 days per week",
                              "Frequently: 3 or more days per week" = "Frequently: 3 or more days per week",
                              "Sometimes: 1-2 days per month but less than weekly" = "Sometimes: 1-2 days per month\nbut less than weekly",
                              "Occasionally: 1-2 days per quarter but less than monthly" = "Occasionally: 1-2 days per quarter\nbut less than monthly",
                              "Rarely: 1-2 days per year or less" = "Rarely: 1-2 days per year or less",
                              "Never" = "Never"),
    improve_onboarding_27 = recode(improve_onboarding_27, 
                                   "I don't know" = "NA",
                                   "I’m not sure"= "NA",  
                                   "more"= "NA",
                                   "N/A"= "NA",
                                   "No idea"= "NA" ,
                                   "Not working in a company"= "NA" ,
                                   "I am quitting after nss"= "NA" ,
                                   "With some experience"= "NA",
                                   "I think just keep doing what they're doing will suffice."= "I am happy"  ,
                                   "It's great. Nothing comes to mind"= "I am happy"  ,
                                   "It’s fine"= "I am happy",
                                   "It’s the best"= "I am happy", 
                                   "There is no nee for any improvement"= "I am happy", 
                                   "None"= "I am happy",
                                   "The best I have encounter, won't change a thing."= "I am happy",
                                   
                                   "Communication"=  "Better documentation of infrastructure and work material",
                                   "More documentation of code"=  "Better documentation of infrastructure and work material",
                                   "Documenting a structure"= "Better documentation of infrastructure and work material",
                                   "Proper documentation of existing softwares"= "Better documentation of infrastructure and work material",
                                   "Proper documentation of infrastructure"= "Better documentation of infrastructure and work material",
                                   "Scripts to set up devices, documentation of help new team members familiarize themselves with the different projects" ="Better documentation of infrastructure and work material",
                                   "set up a tutorial on the stack being used, architecture and how the development cycle goes"= "Better documentation of infrastructure and work material",
                                   
                                   "Roles could be properly defined and documented"= "Clearly defined roles",
                                   "By clearly defining the core functions of the various units/departments"= "Clearly defined roles",
                                   "Company needs to allow the organisational structures to work without interference and clear task to be assigned to new hires"= "Clearly defined roles",
                                   "Roles need to be well spelt out and every necessary document handed before start of work."= "Clearly defined roles",
                                   "Set the expectations on the role earlier"= "Clearly defined roles",
                                   
                                   "Through orientation"= "Better orientation for new workers",
                                   "Could have someone to take me through the code instead of having to understand it myself"= "Better orientation for new workers",
                                   "Assign mentors to newly assigned employees"= "Better orientation for new workers",
                                   "Have orientation sessions before on-boarding"= "Better orientation for new workers",
                                   "Orientation for new employees"= "Better orientation for new workers",
                                   "Providing mentorship and an enabling environment for learning"= "Better orientation for new workers",
                                   "Pair a new developer with developer skilled in his art within the organization and is willing to help"= "Better orientation for new workers",
                                   "Slower rollout. Everything happens too fast, so it doesn't stick"= "Better orientation for new workers",
                                   
                                   "A standardized structure for onboard ding"= "Clearer structure and standardisation",
                                   "Having a standard onboarding process"= "Clearer structure and standardisation",
                                   "By defining a structure in the first place"= "Clearer structure and standardisation",
                                   "By defining a structure in the first place"= "Clearer structure and standardisation",
                                   "Structured Onboarding program"= "Clearer structure and standardisation",
                                   "By establishing an HR Department"= "Clearer structure and standardisation",
                                   "Dedicated staff to handle on boarding process"= "Clearer structure and standardisation",
                                   "By improving the organisational structure to facilitate the work flow"= "Clearer structure and standardisation",
                                   "Improve the organizational structure" = "Clearer structure and standardisation", 
                                   "Let IT help desk raise tickets for each unit responsible for delivering into the process" ="Clearer structure and standardisation",
                                   
                                   "Through IT skills" ="Better/more training",
                                   "More specific training to understand fully the role as opposed to general orientation." ="Better/more training",
                                   "Training in specific fields should be better encouraged" ="Better/more training",
                                   "External trainings" ="Better/more training",
                                   "Developing better internship systems for prospective workers" ="Better/more training",
                                   "More coding workshops" ="Better/more training",
                                   "By applying more practical tests for the individual" ="Better/more training",
                                   
                                   "More employment" ="Employing more people",
                                   "to employee more workers" ="Employing more people",
                                   
                                   "Adequate funds for the right role" ="Other",
                                   "A little interview on the chosen choice of language" ="Other",
                                   "Capacity building" ="Other",
                                   "inclusiveness" ="Other",
                                   "Job positions should be made known" ="Other",
                                   "More focus on developers" ="Other",
                                   "More innovation and creativity" ="Other",
                                   "Retention of talents.." ="Other",
                                   "Observation through hard work" ="Other",
                                   "Better background check" ="Other",
                                   "Need new machines" ="Other",
                                   "More interaction with leads" ="Other"
    ),
    improve_onboarding_27 = fct_relevel(improve_onboarding_27,
                                        "Other",
                                        "Employing more people",
                                        "Clearly defined roles",
                                        "Better documentation of infrastructure and work material",
                                        "Better/more training",
                                        "Better orientation for new workers",
                                        "Clearer structure and standardisation",
                                        "I am happy"))


skills <- rio::import("data/clean/skills_final.csv") %>% 
  select(-V1) %>%  # artifact due to a save as .csv to remove (at least on a Mac?)
  mutate_all(na_if,"") # the column 'level' contains blanks ("") instead of NAs -> transform them to NAs.
# we add characteristics of the respondents to the skills, 
# so that we can differentiate their popularity by occupation or gender:
skills_with_characteristics <- skills %>% left_join(select(Qs, ID, profession_1, gender_35), by = c("id" = "ID"))

number_of_respondents = dim(Qs)[1] # number of respondents (272) used to compute percentages.
number_of_prof_dev = dim(Qs %>% filter(profession_1 == "I am a developer by profession"))[1]
number_of_students = dim(Qs %>% filter(profession_1 == "I am a student who is learning to code"))[1]

```

```{r}

# Themes and Functions:

# Definition of some styling variables for the plots,
# they are used in the theme and plotting functions.
highlight_col = "#d4145a"
text_col = "#777777"
header_col = "#222222"
font_in_viz = "sans" # !!! SHOULD WE CHANGE ????
text_col_in_viz = header_col
main_fontsize = 10 # 'default' base size for the plots
HL_colors = c("#59B0E3","#E3E19A","#6FE38E","#AF8CDE","#DF4B4F")
colours = c( "#610b70", "#88b101", "#eb1c96", "#e98403", "#45454C", "#000001",  "black")
colours_6 = c( "#E04B4F", "#C55F6C", "#AA738A", "#8E87A7", "#739BC5", "#58AFE2")
colours_8 = c( "#E04B4F", "#C55F6C", "#AA738A", "#8E87A7", "#739BC5", "#58AFE2", "#000000", "#000001")
colours_11 = c( "#E04B4F", "#D4545C", "#C75D6A", "#C55F6C", "#AA738A", "#BB6677", "#AF6F84", "#8E87A7", "#739BC5", "#58AFE2", "#000000", "#000001", "pink", "green", "blue", "grey", "red", "steelblue", "orange", "purple") # Assign colors to study fields
# HL_colors_sequential = c(text_col, "#610b70","#88b101","#eb1c96","#e98403","#454545")
geom_text_size = 3.5 # size of text in plots


barplot_theme <- function(){
  # the definition of a custom ggplot theme for our barplots.
  theme_minimal() + 
    theme(
      plot.title = element_text(family = font_in_viz, size = main_fontsize + 2, color = text_col_in_viz),
      # plot.subtitle = element_text(family = font_in_viz, size = main_fontsize),
      axis.title = element_text(family = font_in_viz, size = main_fontsize, color = text_col_in_viz),
      axis.text = element_text(family = font_in_viz, size = main_fontsize, color = text_col_in_viz),
      # plot.caption = element_text(hjust = 0, family = my_font, 
      #                             size = main_fontsize, face= "italic"), #Default is hjust=1 / 0
      plot.title.position = "plot",  # left-align title
      # plot.caption.position =  "plot", 
      panel.grid.minor.x = element_blank(),  # remove useless minor grid
      panel.grid.major.y = element_blank(),  # remove useless major grid in bar charts
      panel.grid.major.x = element_blank(),  # remove useless major grid in bar charts
      axis.title.x = element_blank(), # we do not want an y axis for the barplot, we have labels already
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
      # plot.margin = margin(0.1, 0.1, 0.1, 0.1, unit = "cm")
    )
}

basic_barplot <- function(my_df, a_factor, its_values, labels, title){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = fct_reorder({{a_factor}}, {{its_values}}), 
                             y = {{its_values}})) + 
    # LOLLIPOP:
    geom_point(color = highlight_col, size = 5) +
    geom_segment(aes(x={{a_factor}}, xend={{a_factor}},
                     y=0, yend={{its_values}}),
                 color = highlight_col,
                 size = 1) +
    # # rounding the edges of the bars: 
    # geom_chicklet(radius = grid::unit(1, 'mm'), fill = highlight_col) + 
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.07*max({{its_values}}))),
      size = geom_text_size
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}

#Function for barplot that does not override order
barplot_ordered <- function(my_df, a_factor, its_values, labels, title){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = {{a_factor}},
                             y = {{its_values}})) + 
    # # CHICKLET
    # geom_chicklet(radius = grid::unit(1, 'mm'), fill = highlight_col) +
    # SKINNY LOLLIPOPO:
    geom_point(color = highlight_col, size = 5) +
    geom_segment(aes(x={{a_factor}}, xend={{a_factor}},
                     y=0, yend={{its_values}}),
                 color = highlight_col,
                 size = 1) +
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.07*max({{its_values}}))),
      size = geom_text_size
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}

compute_perc <- function(my_df, a_factor){
  # a small function to count a factor and compute the percentages
  # also return the percentages as char for labelling.
  # USE THIS FUNCTION IF THE SUM OF THE ANSWERS IS 272! (including NAs) -> not working if several possible choices.
  my_df %>%
    count({{a_factor}}) %>%
    drop_na() %>% # always in percentage of the respondents to the question.
    mutate(
      perc = n/sum(n),
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% return()
  
}

compute_perc_popular_skills <- function(the_skills, skillset, number_of_resp){
  # USE THIS FUNCTION for the popular skills (last year + both)
  # NOTE: we use all respondents, not only those who responded. -> some respondents did not enter any skills. -> not optimal, but annoying to correct?
  the_skills %>%
    filter(tool %in% skillset) %>% # select only the languages from all the tools
    filter(level == "Worked with in PAST year" | level == "Both") %>%
    count(tool) %>%
    mutate(
      perc = n/number_of_resp, 
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% 
    arrange(-perc) %>%  
    return()
}

compute_perc_future_skills <- function(the_skills, skillset, number_of_resp){
  # USE THIS FUNCTION for the future, nexy year, skills (next year only)
  # NOTE: we use all respondents, not only those who responded. -> some respondents did not enter any skills. -> not optimal, but annoying to correct?
  the_skills %>%
    filter(tool %in% skillset) %>% # select only the languages from all the tools
    filter(level == "Want to work with NEXT year") %>%
    count(tool) %>%
    mutate(
      perc = n/number_of_resp, 
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% 
    arrange(-perc) %>%  
    return()
}



bubble_plotly_theme <- function(){
  theme_minimal() +
    theme(legend.position="none",
          panel.grid.major.y = element_blank(),
          panel.grid.major.x = element_blank()
    )
}


bubble_plot <- function(df, x_value, y_value, freq, colour_choice, title){
  ggplot(data = df, aes(x={{x_value}}, y={{y_value}}, 
                        text=text)) + # text = text that appears when hover over the bubble, needs to be defined beforehand
    geom_point(aes(size=ifelse({{freq}}==0, NA, {{freq}}),  # bubbles with the size according to how many respondents
                   fill = {{x_value}}), # bubble color according to x value level
               alpha = 0.75, 
               shape = 19, color=NA) + # circles and no border
    scale_size(range = c(1.4, 15)) + # size of the bubbles
    labs( x= "", y = "", size = "Number of respondents", fill = "") +
    scale_fill_manual(values = {{colour_choice}}) +
    scale_colour_manual(values= "white")
  
}


# Trying to optimize the heights of the plots:
#   heights depends on number of "bars"
#   e.g. b3 means all the plots with 3 bars have an fig.height of 1.5
b2 <- 1
b3 <- 1.5
b4 <- 1.8
b6 <- 2.5
b10 <- 3
b18 <- 4

```

# Overview

Hacklab Ghana Developer Census 2020 is the first and most comprehensive survey of people who code in Ghana. In 2020, the Hacklab Foundation fielded a survey covering everything from developers’ favorite technologies to their job preferences. This marks the first of annual surveys to be published. This maiden edition witnessed the participation of 266 developers who took the 20-minute survey earlier between November - December 2020.   

Despite our survey’s reach and capacity for informing valuable conclusions, we acknowledge that our results don’t represent everyone in the Ghana developer community evenly. We have further work to do to make the Hacklab Ghana Developer Census 2020 a more inclusive, diverse platform, and a reflection in the community at large.   

We are committed to building on steps we have taken and improving the coverage, insights and beyond in the coming years to better improve the support and interventions needed to give developers a more enabling ecosystem to thrive. Some of these survey results directly guide those efforts. To address the characteristics of our data, be sure to check out where we summarize results by developer persona (Professional, Student, Hobbyists) or gender, highlight results for underrepresented racial/ethnic groups, or use survey weighting to correct for demographic skew.   

We looked at breakdowns by demographics throughout our analysis and its reflection of the distribution of talents.    

Want to dive into the results yourself? The anonymized results of the survey are [available for download]() under the [Open Database License (ODbL)](https://opendatacommons.org/licenses/odbl/1-0/). We look forward to seeing what you find!  

This maiden edition could not have been successful without the contributions of [Twitter](https://twitter.com/?lang=en) and [CorrelAid](https://correlaid.org/).  


## Key Insights

Here are a few of the top takeaways from this year’s results.  [_add links and a few sentences_]

1.	Gender Representation
2.	Geographical concentration
3.	Most known Languages
4.	Most familiar Frameworks
5.	Communities
6.	Highest level of education
7.	Primary field of study (Business comes 3rd after Computer Engineering/Science and IT )
8.	Overtime & Compensation







<hr>






# Developer Profile

## Type of Developer

### What describes you best?

The two biggest subgroups among the respondents are professional developers as well as students. Additionally, there are respondents coding as a part of their work, coding as a hobby, as well as former developers. 

```{r, fig.align='center', out.width = '100%', fig.height = b6}

basic_barplot(my_df = compute_perc(Qs, profession_1), 
              a_factor = profession_1, its_values = perc, labels = perc_label, 
              title = "")
```

### Do you code as a hobby? 

Most of the respondents code as a hobby. Interestingly, professional developers seem to code for a hobby less often than students which may be related to their reduced time availability. 

::::: {.panelset}

::: {.panel}

##### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b2}

basic_barplot(my_df = compute_perc(Qs, hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```


:::

::: {.panel}

##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height = b2}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

##### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b2}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::

## Employment

```{r}

employment_data <- Qs %>%
  filter(employment_3 != "I prefer not to say") %>%
  mutate(
    employment_3 = factor(employment_3, 
                          levels(employment_3)[c(4,2,3,6,5,1)])
  )
```

Most of the respondents full-time employees. There is also a great share of students. Sadly, there is also a significant share of unemployed developers looking for work. 

When it comes to gender differences, there is a greater share of students among the female respondents. In addition, women seem to be self-employed less often when compared with men. 

::::: {.panelset}

::: {.panel}

##### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b6}

basic_barplot(my_df = compute_perc(employment_data, employment_3), 
              a_factor = employment_3, its_values = perc, labels = perc_label,
              title = "")

```


:::

::: {.panel}

##### Men only {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b6}

barplot_ordered(my_df = compute_perc(employment_data %>% filter(gender_35 == "Man"), employment_3), 
                a_factor = employment_3, its_values = perc, labels = perc_label,
                title = "")

```

:::

::: {.panel}

##### Women only {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b6}

barplot_ordered(my_df = compute_perc(employment_data %>% filter(gender_35 == "Woman"), employment_3), 
                a_factor = employment_3, its_values = perc, labels = perc_label,
                title = "")

```

:::

::::



## Geography 

### Region

Most respondents come from the greater Accra region. However, the extent of concentration in Accra seems to be larger for professional developers than for students. 

::::: {.panelset}

::: {.panel}

##### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b10}
region_data <- Qs %>%
  filter(region_4 != "Not in Ghana")

region_data$region_4 <- recode(region_data$region_4,
                               "Eastern Region, Ghana" = "Eastern Region", 
                               "Northern Region, Ghana" = "Northern Region", 
                               "Central Region, Ghana" = "Central Region", 
                               "Western Region, Ghana"= "Western Region"
)

basic_barplot(my_df = compute_perc(region_data, region_4), 
              a_factor = region_4, its_values = perc, labels = perc_label,
              title = "")

```


:::

::: {.panel}

##### Professional Developers {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b10}

basic_barplot(my_df = compute_perc(region_data %>% filter(profession_1 == "I am a developer by profession"), region_4), 
              a_factor = region_4, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

#####  Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b10}

basic_barplot(my_df = compute_perc(region_data %>% filter(profession_1 == "I am a student who is learning to code"), region_4), 
              a_factor = region_4, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::




### City

Zoom in to get a more detailed picture. 

Hover over the individual markers to see the professional status of the respondents. 

```{r, fig.align= 'center', out.width= '100%', fig.height = 6.5}

# the geocoding was done separately in data_cleaning/Geocoding.R
# (to not lose time everytime we knitr)
# we load the result:
data_city_geo <- rio::import("data/clean/data_city_geo.rds") 

#Map for the cities
violet_icon <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/pointhi/leaflet-color-markers/master/img/marker-icon-violet.png", 
  iconWidth = 24,
  iconHeight = 32) 

leaflet(data = data_city_geo) %>%
  addTiles() %>% 
  addMarkers(~ geocode$lon, ~ geocode$lat, 
             clusterOptions = markerClusterOptions(), 
             icon = violet_icon, label = ~profession_1)
```



## Demographics


### Age

Most respondents are between 20 and 30 years old. 

While the students are very young in most cases, the professional developers display a little more variance in their ages. However, we have a very young sample at hand.

```{r}
Qs$age_range_6 <- factor(Qs$age_range_6, levels(Qs$age_range_6)[c(6:1)])

```

::::: {.panelset}

::: {.panel}

#####  All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b6}

barplot_ordered(my_df = compute_perc(Qs, age_range_6), 
                a_factor = age_range_6, its_values = perc, labels = perc_label, 
                title = "")

```


:::

::: {.panel}

##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height = b6}


barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), age_range_6), 
                a_factor = age_range_6, its_values = perc, labels = perc_label,
                title = "")

```

:::

::: {.panel}

##### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b6}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), age_range_6), 
                a_factor = age_range_6, its_values = perc, labels = perc_label,
                title = "")

```

:::

::::



### Gender 

#### What Gender do you identify with?

There is a  greater share of women among the students than among the professional developers. 

Ghana's developer community may become more representative in the upcoming years. 

```{r}

#Removing NA

gender_data <- Qs %>%
  filter(gender_35 != "NA")
```


::::: {.panelset}

::: {.panel}

##### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b2}

barplot_ordered(my_df = compute_perc(gender_data, gender_35), 
                a_factor = gender_35, its_values = perc, labels = perc_label, 
                title = "")

```


:::

::: {.panel}

##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height = b2}


barplot_ordered(my_df = compute_perc(gender_data %>% filter(profession_1 == "I am a developer by profession"), gender_35), 
                a_factor = gender_35, its_values = perc, labels = perc_label,
                title = "")

```

:::

::: {.panel}

##### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height = b2}

barplot_ordered(my_df = compute_perc(gender_data %>% filter(profession_1 == "I am a student who is learning to code"), gender_35), 
                a_factor = gender_35, its_values = perc, labels = perc_label,
                title = "")

```

:::

::::


#### Do you identify as transgender? 

```{r}
transgender_data <- Qs %>%
  filter(transgender_36 != "NA")
```


```{r, fig.align='center', out.width = '100%', fig.height = b2}

basic_barplot(my_df = compute_perc(transgender_data, transgender_36), 
              a_factor = transgender_36, its_values = perc, labels = perc_label,
              title = "")

```


### Sexual orientation

```{r}
sexual_orientation_data <- Qs %>%
  filter(sexual_orientation_37 != "NA" & sexual_orientation_37 != "Prefer not to say")

```

```{r, fig.align='center', out.width = '100%', fig.height = b4}

basic_barplot(my_df = compute_perc(sexual_orientation_data, sexual_orientation_37), 
              a_factor = sexual_orientation_37, its_values = perc, labels = perc_label,
              title = "")

```

### Ethnicity 

```{r, fig.align='center', out.width = '100%', fig.height = b2}

basic_barplot(my_df = compute_perc(Qs %>% filter(ethnicity_38 != "NA" & ethnicity_38 != "I prefer not to answer"), ethnicity_38), 
              a_factor = ethnicity_38, its_values = perc, labels = perc_label,
              title = "")

```

### Do you have any dependents you care for? 

Among our respondents, women seem to have dependents they care for less often than men do. This may partly be due to the fact, that more women are still in their studies and that the women in our sample may be younger than the men. 

```{r}
dependents_data <- Qs %>%
  filter(dependents_41 != "I prefer not to say" & dependents_41 != "NA")
```

::::: {.panelset}

::: {.panel}

##### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=1}

basic_barplot(my_df = compute_perc(dependents_data, dependents_41), 
              a_factor = dependents_41, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

##### Only Men {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=1}

basic_barplot(my_df = compute_perc(dependents_data %>% filter(gender_35 == "Man"), dependents_41), 
              a_factor = dependents_41, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

##### Only Women {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=1}

basic_barplot(my_df = compute_perc(dependents_data %>% filter(gender_35 == "Woman"), dependents_41), 
              a_factor = dependents_41, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::








<hr>










# Technology and Tech Culture 

## Most Popular Technologies

These are the technologies with which the respondents have done extensive development work in over the past year.  


### Programming, scripting and markup languages

_HTML/CSS_ is the most used language, followed by _JavaScript_, _Python_ and _SQL_. Professional developers are more likely to use _JavaScript_ and _SQL_. _C++_ and _C_ are still widely used to teach coding to students.    


```{r}

languages = c(
  "C",			
  "C#",			
  "C++",		
  "HTML/CSS",			
  "Java",			
  "JavaScript",			
  'Kotlin',			
  "PHP",			
  "Python",			
  "Ruby",			
  "SQL",			
  "Swift",			
  "TypeScript"
)


```
::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics, 
                                                  skillset = languages,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = languages,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = languages,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


### Platforms

A comment here...  

```{r}

platforms = c(
  "Android",			
  "Arduino",			
  "AWS",			
  "Docker",		
  "Google Cloud Platform",			
  "Heroku",			
  "IBM Cloud or Watson",			
  "iOS",			
  "Linux",			
  "MacOS",		
  "Microsoft Azure",		
  "Raspberry Pi", 	
  "Windows",			
  "WordPress"
)

```
::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics, 
                                                  skillset = platforms,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = platforms,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = platforms,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


### Web frameworks

_React.js_ is the most popular web framework.   

```{r}

web_frameworks = c(
  "Angular",			
  "ASP.NET",		
  "Django",			
  "Express",		
  "jQuery",			
  "Laravel",
  "React.js",
  "Ruby on Rails",
  "Vue.js",
  "Spring"
)

```
::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics, 
                                                  skillset = web_frameworks,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = web_frameworks,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = web_frameworks,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


### Other frameworks, libraries and tools

_Node.js_ is also a popular back-end environment.    

```{r}

other_frameworks = c(
  ".NET",			
  ".NET Core",			
  "Flutter",			
  "Node.js",			
  "Puppet",			
  "React Native",			
  "Deno.js"
)

```
::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2.5}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics, 
                                                  skillset = other_frameworks,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2.5}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = other_frameworks,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2.5}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = other_frameworks,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


### Collaborative tools

_GitHub_ is by far the most popular version control tool. _Slack_ is the most used business communication platform.  

```{r}

collaborative_tools = c(
  "GitHub",			
  "GitLab",			
  "Facebook Workplace",			
  "Slack",			
  "Microsoft Teams",			
  "Microsoft Azure",			
  "Trello",			
  "Google Suite (Docs, Meet, etc)",			
  "Stack Overflow for Teams",			
  "Jira"
)

```
::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics, 
                                                  skillset = collaborative_tools,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = collaborative_tools,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = collaborative_tools,
                                                  number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::



## Wanted Technologies 

These are the technologies that the respondents are currently not using but want to work in over the next year.  

### Programming, scripting and markup languages

_TypeScript_, followed by _Python_ and _Swift_ are the technologies that professional developers do not use but want to learn. Students are interested in "the classics": _Python_, _SQ_, and _JavaScript_.


::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics, 
                                                 skillset = languages,
                                                 number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")
```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                   filter(profession_1 == "I am a developer by profession"), 
                                                 skillset = languages,
                                                 number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                   filter(profession_1 == "I am a student who is learning to code"), 
                                                 skillset = languages,
                                                 number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


### Platforms

Respondents want to learn working with cloud platforms, in particular _AWS_ for the professional developers and _Microsoft Azure_ for the students. Almost 20% of professional developers indicated wanting to learn _Docker_. _iOS_ is the mobile developing system that most want to work with nextt year.        

::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics, 
                                                 skillset = platforms,
                                                 number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                   filter(profession_1 == "I am a developer by profession"), 
                                                 skillset = platforms,
                                                 number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                   filter(profession_1 == "I am a student who is learning to code"), 
                                                 skillset = platforms,
                                                 number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


### Web frameworks

If _React.js_ was [the most popular web framework](web-frameworks), it seems that _Django_ and _Vue.js_ are attractive.      

```{r}

web_frameworks = c(
  "Angular",			
  "ASP.NET",		
  "Django",			
  "Express",		
  "jQuery",			
  "Laravel",
  "React.js",
  "Ruby on Rails",
  "Vue.js",
  "Spring"
)

```
::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics, 
                                                 skillset = web_frameworks,
                                                 number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                   filter(profession_1 == "I am a developer by profession"), 
                                                 skillset = web_frameworks,
                                                 number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                   filter(profession_1 == "I am a student who is learning to code"), 
                                                 skillset = web_frameworks,
                                                 number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


### Other frameworks, libraries and tools

A quarter of the respondents indicated wanting to work with _Flutter_ next year. Many also want to work with _React Native_ and _Node.js_.  

```{r}

other_frameworks = c(
  ".NET",			
  ".NET Core",			
  "Flutter",			
  "Node.js",			
  "Puppet",			
  "React Native",			
  "Deno.js"
)

```
::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2.5}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics, 
                                                 skillset = other_frameworks,
                                                 number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2.5}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                   filter(profession_1 == "I am a developer by profession"), 
                                                 skillset = other_frameworks,
                                                 number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2.5}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                   filter(profession_1 == "I am a student who is learning to code"), 
                                                 skillset = other_frameworks,
                                                 number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


### Collaborative tools

_Microsoft Azure_ and _Stack Overflow for Teams_ are tools who might become more popular. Many students also indicate wanting to learn _GitHub_.  

```{r}

collaborative_tools = c(
  "GitHub",			
  "GitLab",			
  "Facebook Workplace",			
  "Slack",			
  "Microsoft Teams",			
  "Microsoft Azure",			
  "Trello",			
  "Google Suite (Docs, Meet, etc)",			
  "Stack Overflow for Teams",			
  "Jira"
)

```
::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics, 
                                                 skillset = collaborative_tools,
                                                 number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                   filter(profession_1 == "I am a developer by profession"), 
                                                 skillset = collaborative_tools,
                                                 number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_future_skills(the_skills = skills_with_characteristics %>% 
                                                   filter(profession_1 == "I am a student who is learning to code"), 
                                                 skillset = collaborative_tools,
                                                 number_of_resp = number_of_students), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::



## Operating System

### What is the primary operating system in which you work?

Windows in the most used operating system. Professional developers use MacOS and Linux more often than other respondents. Students are overwhelmingly on Windows.  

::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=2}

basic_barplot(my_df = compute_perc(Qs, prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=2}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=2}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::


### In which operating system would you rather work?

MacOS, Linux and Windows are practically tied as preferred systems. Professional developers have a preference for MacOS, then Linux. Students prefer Windows.  

::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2}

basic_barplot(my_df = compute_perc(Qs, rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```
:::
:::::



## New Technologies

### How frequently do you learn a new language or framework?

::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2.5}

barplot_ordered(my_df = compute_perc(Qs, new_tool_7), 
                a_factor = new_tool_7, its_values = perc, labels = perc_label, 
                title = "") + scale_x_discrete(limits=rev)

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2.5}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), new_tool_7), 
                a_factor = new_tool_7, its_values = perc, labels = perc_label, 
                title = "") + scale_x_discrete(limits=rev)

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2.5}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), new_tool_7), 
                a_factor = new_tool_7, its_values = perc, labels = perc_label, 
                title = "") + scale_x_discrete(limits=rev)

```
:::
:::::


### What level of influence do you, personally, have over new technology purchases at your organization?

Unsurprisingly, professional developers are more likely to have influence over new technology purchases.  

::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2}

barplot_ordered(my_df = compute_perc(Qs, purch_influence_16), 
                a_factor = purch_influence_16, its_values = perc, labels = perc_label, 
                title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% 
                                       filter(profession_1 == "I am a developer by profession"),
                                     purch_influence_16), 
                a_factor = purch_influence_16, its_values = perc, labels = perc_label, 
                title = "")

```
:::
:::::

### When buying a new tool or software, how do you discover and research available solutions? Select all that apply.

Most of the respondents declared relying on free trials and colleagues.  

```{r, fig.align='center', out.width = '100%', fig.height=3}

# multiple choice question: we separate the different answers (separate_rows),
#   and compute the % according to the total number of respondents
perc_solution_research_17 <- Qs %>% 
  select(ID, profession_1, solution_research_17) %>% 
  separate_rows(solution_research_17, sep = ";") %>% 
  # drop_na(any_of("solution_research_17")) %>% 
  filter(solution_research_17 != "") %>% 
  mutate(
    solution_research_17 = ifelse(solution_research_17 == "YouTube Videos", "YouTube", solution_research_17), # group those
    solution_research_17 = ifelse(solution_research_17 == "YouTube ", "YouTube", solution_research_17) # group those
  ) %>%
  count(solution_research_17) %>% 
  mutate(
    perc = n/number_of_respondents,  # % to total number of respondents -> some resp. selected several choices
    perc_label = paste0(as.character(round(perc*100,1)), "%")
  ) %>% 
  arrange(-perc) 

basic_barplot(my_df = perc_solution_research_17, 
              a_factor = solution_research_17, its_values = perc, labels = perc_label, 
              title = "")



```








<hr>














# Education, Work and Career

## Education

### Highest level of formal education

Most of the survey participants have at least a secondary high-school degree whereas half of the students have a secondary high school degree and every third professional developer has a bachelor's degree

::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, highest_edu_18), 
                a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
                title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), highest_edu_18), 
                a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
                title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), highest_edu_18), 
                a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
                title = "")

```
:::
:::::


### Primary field of study

Most participants study or studied Computer Science or Computer Engineering, followed by Information Technology and other Engineering fields as well as Mathematics/Statistics and Business.

::::: {.panelset}
::: {.panel}

##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height = b18}

basic_barplot(my_df = compute_perc(Qs, prim_study_19), 
                a_factor = prim_study_19, its_values = perc, labels = perc_label, 
                title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height = b18}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), prim_study_19), 
                a_factor = prim_study_19, its_values = perc, labels = perc_label, 
                title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height = b18}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), prim_study_19), 
                a_factor = prim_study_19, its_values = perc, labels = perc_label, 
                title = "")

```
:::
:::::

#### Does the primary field of study matter for the salary as a professional developer?

```{r, fig.align='center', out.width = '100%', fig.height=3, warning=FALSE, message=FALSE}

# Responsive Bubble Chart for Relationship between Job Satisfaction and Overtime Work

prim_study_19_28_prof <- Qs %>%
  filter(profession_1 == "I am a developer by profession") %>%
  mutate(monthly_salary_28_7 = recode(monthly_salary_28,
                                      "Greater than GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 20,000 - GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 15,000 - GHS 20,000"= "Greater than GHS 15,000",
                                      "GHS 10,000 - GHS 15,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 8,000 - GHS 10,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 6,000 - GHS 8,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 5,000 - GHS 6,000"= "GHS 3,500 - GHS 8,000",
                                      "GHS 4,000 - GHS 5,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 3,500 - GHS 4,000"= "GHS 3,500 - GHS 8,000"))  


runningcounts.df <- as.data.frame(table(prim_study_19_28_prof$prim_study_19, prim_study_19_28_prof$monthly_salary_28_7))

relation_ordinal_Q19_28 <- runningcounts.df %>% rename(monthly_salary_28 = Var2,
                                                       prim_study_19 = Var1)

levels(relation_ordinal_Q19_28$monthly_salary_28) <- gsub(" - ", "\n", levels(relation_ordinal_Q19_28$monthly_salary_28))
levels(relation_ordinal_Q19_28$prim_study_19) <- sub(" ", "\n", levels(relation_ordinal_Q19_28$prim_study_19))
levels(relation_ordinal_Q19_28$prim_study_19) <- sub("/", "\n", levels(relation_ordinal_Q19_28$prim_study_19))

relation_ordinal_Q19_28 <- relation_ordinal_Q19_28 %>%
  # prepare text for tooltip
  mutate(text = paste(Freq, "respondent/s")) %>%
  filter(Freq > 0)

# # Heatmap as an alternative
# ggHeatmap(relation_ordinal_Q19_28,
#           aes(x = monthly_salary_28,
#               y = prim_study_19,
#               fill = Freq),
#           addlabel=FALSE,
#           interactive=TRUE,
#           tooltip = text)


names(colours_8) <- levels(relation_ordinal_Q19_28$monthly_salary_28)
colScale <- scale_colour_manual(name = "monthly_salary_28",values = colours_8)

bubble_Q19_28_resp = ggplot(data = relation_ordinal_Q19_28, 
                            aes(x=monthly_salary_28, y=prim_study_19, color=monthly_salary_28)) +
  geom_point_interactive(aes(size=ifelse(Freq==0, NA, Freq),
                             tooltip = text)) +
  colScale +
  labs(x = "", y = "") +
  bubble_plotly_theme()

girafe(ggobj = bubble_Q19_28_resp, 
       fonts = list(sans = "Arial"),
       width_svg = 7, 
       height_svg = 4,
       options = list(
         opts_tooltip(use_fill = TRUE)))

```


#### How do Study Field, Degree and Salary relate to each other?

```{r, fig.align='center', out.width = '100%', fig.height=3, warning=FALSE}
prim_study_19_28_prof <- Qs %>%
  filter(profession_1 == "I am a developer by profession",
         highest_edu_18 == "Bachelor" | 
           highest_edu_18 == "Master" |
           highest_edu_18 == "Secondary High School" |
           highest_edu_18 == "Higher National Diploma") %>%
  filter(profession_1 == "I am a developer by profession") %>%
  mutate(monthly_salary_28_7 = recode(monthly_salary_28,
                                      "Greater than GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 20,000 - GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 15,000 - GHS 20,000"= "Greater than GHS 15,000",
                                      "GHS 10,000 - GHS 15,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 8,000 - GHS 10,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 6,000 - GHS 8,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 5,000 - GHS 6,000"= "GHS 3,500 - GHS 8,000",
                                      "GHS 4,000 - GHS 5,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 3,500 - GHS 4,000"= "GHS 3,500 - GHS 8,000"))
# drop Education Levels that have 0 respondents
prim_study_19_28_prof$highest_edu_18<- droplevels(prim_study_19_28_prof$highest_edu_18)

runningcounts.df <- as.data.frame(table(prim_study_19_28_prof$prim_study_19, prim_study_19_28_prof$monthly_salary_28_7, prim_study_19_28_prof$highest_edu_18))

prim_study_19_28_prof <- runningcounts.df %>% rename(monthly_salary_28 = Var2,
                                                     prim_study_19 = Var1,
                                                     highest_edu_18 = Var3)

# prepare text for tooltip
prim_study_19_28_prof <- prim_study_19_28_prof %>%
  mutate(text = paste("study field: ", prim_study_19,"\n", Freq, "respondent/s"))


names(colours_11) <- levels(prim_study_19_28_prof$prim_study_19)
colScale <- scale_colour_manual(name = "prim_study_19",values = colours_11)


levels(prim_study_19_28_prof$monthly_salary_28) <- gsub(" - ", "\n", levels(prim_study_19_28_prof$monthly_salary_28))
levels(prim_study_19_28_prof$highest_edu_18) <- gsub(" ", "\n", levels(prim_study_19_28_prof$highest_edu_18))

bubble_Q19_28_resp = ggplot(data = prim_study_19_28_prof, 
                            aes(x=highest_edu_18, y=monthly_salary_28, color= prim_study_19)) +
  geom_jitter_interactive(width = 0.3, height = 0.3, aes(size=ifelse(Freq==0, NA, Freq),
                                                         tooltip = text)) +
  colScale +
  labs(x = "", y = "") +
  bubble_plotly_theme()+
  theme_bw() +
  theme(legend.position="none")


girafe(ggobj = bubble_Q19_28_resp, 
       fonts = list(sans = "Arial"),
       width_svg = 6, 
       height_svg = 4,
       options = list(
         opts_tooltip(use_fill = TRUE)))

```

### How important is a formal education, such as a university degree in computer science, to your career?

Half of the respondents find a formal education very or even critically important for their career.

::::: {.panelset}
::: {.panel}
##### All Respondents {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height= 2}

barplot_ordered(my_df = compute_perc(Qs, edu_importance_20), 
                a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
                title = "")

```
:::
::: {.panel}
##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height= 2}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), edu_importance_20), 
                a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
                title = "")

```
:::
::: {.panel}
##### Students {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height= 2}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), edu_importance_20), 
                a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
                title = "")

```
:::
:::::


#### If you could go back and change your educational path (but end up in the same career), what would you change?

```{r, fig.align='center', out.width = '100%', fig.height=3}

compute_perc(Qs, change_edu_21 )

```



## Work

### Job satisfaction

A little bit than more of 50% of the respondents are slightly satisfied or satisfied with their current jobs.
Whereas 20% are slightly dissatisfied or dissatisfied.

```{r, fig.align='center', out.width = '100%', fig.height=2}

barplot_ordered(my_df = compute_perc(Qs, job_satisfaction_22 ), 
                a_factor = job_satisfaction_22 , its_values = perc, labels = perc_label, 
                title = "Q22. How satisfied are you with your current job?")


```

#### Relationship between job satisfaction and overtime work

Interestingly, people who work more than 3 times a week overtime are satisfied with their job. There seems no relationship that the more overtime the less job satisfaction. The tendency is rather the opposite. 

```{r, include = FALSE, fig.align='center', out.width = '100%', fig.height=3}

# Bubble chart: Relationship between job satisfaction and overtime work


# bubble_Q22_25<-bubble_plot(df = relation_ordinal_Q22_25,
#                            x_value = overtime_work_25,
#                            y_value = job_satisfaction_22, freq = Freq, colour_choice = colours, title = "") + bubble_plotly_theme()
#
# # ggplotly(bubble_Q22_25, tooltip="text", width = 800, height = 600)
```

```{r, fig.align='center', out.width = '100%', fig.height=3}

# Responsive Bubble Chart for Relationship between Job Satisfaction and Overtime Work

runningcounts.df <- as.data.frame(table(Qs$job_satisfaction_22, Qs$overtime_work_25))

relation_ordinal_Q22_25 <- runningcounts.df %>% rename(overtime_work_25 = Var2,
                                                       job_satisfaction_22 = Var1) %>%
  mutate(overtime_work_25 = recode(overtime_work_25,
                                   "Often: 1-2 days per week" = "1-2 days per week",
                                   "Sometimes: 1-2 days per month but less than weekly" = "1-2 days per month",
                                   "Rarely: 1-2 days per year or less" = "<1-2 days per year",
                                   "Occasionally: 1-2 days per quarter but less than monthly" = "1-2 days per quarter",
                                   "Frequently: 3 or more days per week" = ">3 days per week"))

levels(relation_ordinal_Q22_25$overtime_work_25) <- gsub(" ", "\n", levels(relation_ordinal_Q22_25$overtime_work_25))
levels(relation_ordinal_Q22_25$job_satisfaction_22) <- gsub(" ", "\n", levels(relation_ordinal_Q22_25$job_satisfaction_22))



relation_ordinal_Q22_25 <- relation_ordinal_Q22_25 %>%
  # prepare text for tooltip
  mutate(text = paste(Freq, "respondent/s"))


names(colours) <- levels(relation_ordinal_Q22_25$overtime_work_25)
colScale <- scale_colour_manual(name = "overtime_work_25",values = colours_6)

bubble_Q22_25_resp = ggplot(data = relation_ordinal_Q22_25, aes(x=overtime_work_25, y=job_satisfaction_22, color=overtime_work_25)) +
  geom_point_interactive(aes(size = Freq,
                             tooltip = text)) +
  colScale +
  labs(x = "", y = "") +
  bubble_plotly_theme()


girafe(ggobj = bubble_Q22_25_resp, 
       fonts = list(sans = "Arial"),
       width_svg = 6, 
       height_svg = 4,
       options = list(
         opts_tooltip(use_fill = TRUE)))
```

### Approximately how many people are employed by the company or organization you currently work for?

Half of the respondents work in a small company with less than 10 workers. Only 10% of the respondents work in a big company with more than 100 workers.

```{r, fig.align='center', out.width = '100%', fig.height= b6}

#Removing NA

company_size_23_NA_remov <- Qs %>%
  filter(company_size_23 != "NA")

barplot_ordered(my_df = compute_perc(company_size_23_NA_remov, company_size_23 ), 
                a_factor = company_size_23 , its_values = perc, labels = perc_label, 
                title = "")
```

### On average, how many hours per week do you work?

Working hours are quite evenly scattered across the scale from 0-80 hours a week.

```{r, fig.align='center', out.width = '100%', fig.height=3}

ggplot(compute_perc(Qs,weekly_work_hours_24), aes(x=weekly_work_hours_24)) + 
  geom_histogram(color = "white", fill= highlight_col, binwidth = 10) + 
  labs(x="Weekly working hours", y = "Number of Respondents") +
  scale_x_continuous(breaks = seq(0, 120, by=10)) +
  theme_minimal() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))


```


### How often do you work overtime or beyond the formal time expectation of your job?

Around 50% of the respondents work overtime on 3 or more days in a week.

```{r, fig.align='center', out.width = '100%', fig.height= 2.5}

barplot_ordered(my_df = compute_perc(Qs, overtime_work_25 ), 
                a_factor = overtime_work_25 , its_values = perc, labels = perc_label, 
                title = "")

```

#### Relationship between Overtime Work and Salary of Professional Developers

```{r, fig.align='center', out.width = '100%', fig.height=3, warnings = FALSE}
overtime_work_25_28_prof <- Qs %>%
  filter(profession_1 == "I am a developer by profession") %>%
  mutate(monthly_salary_28_7 = recode(monthly_salary_28,
                                      "Greater than GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 20,000 - GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 15,000 - GHS 20,000"= "Greater than GHS 15,000",
                                      "GHS 10,000 - GHS 15,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 8,000 - GHS 10,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 6,000 - GHS 8,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 5,000 - GHS 6,000"= "GHS 3,500 - GHS 8,000",
                                      "GHS 4,000 - GHS 5,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 3,500 - GHS 4,000"= "GHS 3,500 - GHS 8,000"))  


runningcounts.df <- as.data.frame(table(overtime_work_25_28_prof$overtime_work_25, overtime_work_25_28_prof$monthly_salary_28_7))

relation_ordinal_Q25_28 <- runningcounts.df %>% rename(monthly_salary_28 = Var2,
                                                       overtime_work_25 = Var1)

levels(relation_ordinal_Q25_28$monthly_salary_28) <- gsub(" - ", "\n", levels(relation_ordinal_Q25_28$monthly_salary_28))
levels(relation_ordinal_Q25_28$overtime_work_25) <- sub("days", "days \n", levels(relation_ordinal_Q25_28$overtime_work_25))


relation_ordinal_Q25_28 <- relation_ordinal_Q25_28 %>%
  # prepare text for tooltip
  mutate(text = paste(Freq, "respondent/s")) 


# Heatmap as an alternative
# ggHeatmap(relation_ordinal_Q23_28,
#           aes(x = company_size_23,
#               y = monthly_salary_28,
#               fill = Freq),
#           addlabel=FALSE,
#           interactive=TRUE,
#           tooltip = text)


names(colours_6) <- levels(relation_ordinal_Q25_28$overtime_work_25)
colScale <- scale_colour_manual(name = "overtime_work_25",values = colours_6)

bubble_Q25_28_resp = ggplot(data = relation_ordinal_Q25_28, aes(x=overtime_work_25, y=monthly_salary_28, color=overtime_work_25)) +
  geom_point_interactive(aes(size=ifelse(Freq==0, NA, Freq),
                             tooltip = text)) +
  colScale +
  labs(x = "", y = "") +
  bubble_plotly_theme()



girafe(ggobj = bubble_Q25_28_resp, 
       fonts = list(sans = "Arial"),
       width_svg = 6, 
       height_svg = 4,
       options = list(
         opts_tooltip(use_fill = TRUE)))

```


### Do you think your company has a good on-boarding process?


```{r, fig.align='center', out.width = '100%', fig.height = b2}

basic_barplot(my_df = compute_perc(Qs, company_onboarding_process_26 ), 
              a_factor = company_onboarding_process_26 , its_values = perc, labels = perc_label, 
              title = "")

```

### How could on-boarding at your company be improved?

```{r, fig.align='center', out.width = '100%', fig.height=3}

improve_onboarding_27_filt <- Qs %>%
  filter(improve_onboarding_27 != "NA")

barplot_ordered(my_df = compute_perc(improve_onboarding_27_filt, improve_onboarding_27), 
                a_factor = improve_onboarding_27 , its_values = perc, labels = perc_label, 
                title = "")

```

### How much is your monthly salary in Ghana Cedis?

50% of the respondents earn less than 2000 GHS per month. 

```{r}
monthly_salary_data <- Qs %>%
  filter(monthly_salary_28 != "NA")

```

::::: {.panelset}

::: {.panel}

##### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, monthly_salary_28 ), 
                a_factor = monthly_salary_28 , its_values = perc, labels = perc_label, 
                title = "")

```

:::

::: {.panel}

##### Only Women {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(monthly_salary_data %>% filter(gender_35 == "Woman"), monthly_salary_28), 
                a_factor = monthly_salary_28, its_values = perc, labels = perc_label,
                title = "")

```
:::

::: {.panel}

##### Only Men {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(monthly_salary_data %>% filter(gender_35 == "Man"), monthly_salary_28), 
                a_factor = monthly_salary_28, its_values = perc, labels = perc_label,
                title = "")

```

:::

::::

#### Relationship between Size of company and Salary

The student respondents don't earn more than 2000 GHS but do work in various sizes of companies. Professional Developers earn more salary and the most developers work in companies with less than 100 employees.There seems no clear relationship between the company size and the salary. 

::::: {.panelset}

::: {.panel}

##### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3, warning=FALSE}
company_size_23_NA_remov <- Qs  %>%
  mutate(company_size_23_binned = recode(company_size_23, 
                                         "I don't know" = "NA",
                                         "I’m not sure"= "NA",  
                                         "more"= "NA",
                                         "N/A"= "NA"),
         monthly_salary_28_6 = recode(monthly_salary_28,
                                      "Greater than GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 20,000 - GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 15,000 - GHS 20,000"= "Greater than GHS 15,000",
                                      "GHS 10,000 - GHS 15,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 8,000 - GHS 10,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 6,000 - GHS 8,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 5,000 - GHS 6,000"= "GHS 3,500 - GHS 8,000",
                                      "GHS 4,000 - GHS 5,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 3,500 - GHS 4,000"= "GHS 3,500 - GHS 8,000")) %>%
  filter(company_size_23_binned != "NA") 

company_size_23_NA_remov$company_size_23_binned<- droplevels(company_size_23_NA_remov$company_size_23_binned)


runningcounts.df <- as.data.frame(table(company_size_23_NA_remov$company_size_23_binned, company_size_23_NA_remov$monthly_salary_28_6))

relation_ordinal_Q23_28 <- runningcounts.df %>% rename(monthly_salary_28 = Var2,
                                                       company_size_23 = Var1)

levels(relation_ordinal_Q23_28$monthly_salary_28) <- gsub(" - ", "\n", levels(relation_ordinal_Q23_28$monthly_salary_28))




relation_ordinal_Q23_28 <- relation_ordinal_Q23_28 %>%
  # prepare text for tooltip
  mutate(text = paste(Freq, "respondent/s")) 


# Heatmap as an alternative
# ggHeatmap(relation_ordinal_Q23_28,
#           aes(x = company_size_23,
#               y = monthly_salary_28,
#               fill = Freq),
#           addlabel=FALSE,
#           interactive=TRUE,
#           tooltip = text)


names(colours_6) <- levels(relation_ordinal_Q23_28$company_size_23)
colScale <- scale_colour_manual(name = "company_size_23",values = colours_6)

bubble_Q23_28_resp = ggplot(data = relation_ordinal_Q23_28, aes(x=company_size_23, y=monthly_salary_28, color=company_size_23)) +
  geom_point_interactive(aes(size=ifelse(Freq==0, NA, Freq),
                             tooltip = text)) +
  colScale +
  labs(x = "", y = "") +
  bubble_plotly_theme()



girafe(ggobj = bubble_Q23_28_resp, 
       fonts = list(sans = "Arial"),
       width_svg = 6, 
       height_svg = 4,
       options = list(
         opts_tooltip(use_fill = TRUE)))



# plot responsive plotly
# bubble_Q23_28<-bubble_plot(df = relation_ordinal_Q23_28, x_value = company_size_23, y_value = monthly_salary_28, freq = Freq, colour_choice = colours, title = "") + bubble_plotly_theme()
# ggplotly(bubble_Q23_28, tooltip="text", width = 800, height = 600)
```

:::

::: {.panel}

##### Only Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}
company_size_23_NA_remov <- Qs  %>%
  filter(profession_1 == "I am a developer by profession") %>%
  mutate(company_size_23_binned = recode(company_size_23, 
                                         "I don't know" = "NA",
                                         "I’m not sure"= "NA",  
                                         "more"= "NA",
                                         "N/A"= "NA"),
         monthly_salary_28_6 = recode(monthly_salary_28,
                                      "Greater than GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 20,000 - GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 15,000 - GHS 20,000"= "Greater than GHS 15,000",
                                      "GHS 10,000 - GHS 15,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 8,000 - GHS 10,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 6,000 - GHS 8,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 5,000 - GHS 6,000"= "GHS 3,500 - GHS 8,000",
                                      "GHS 4,000 - GHS 5,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 3,500 - GHS 4,000"= "GHS 3,500 - GHS 8,000")) %>%
  filter(company_size_23_binned != "NA") 

company_size_23_NA_remov$company_size_23_binned<- droplevels(company_size_23_NA_remov$company_size_23_binned)


runningcounts.df <- as.data.frame(table(company_size_23_NA_remov$company_size_23_binned, company_size_23_NA_remov$monthly_salary_28_6))

relation_ordinal_Q23_28 <- runningcounts.df %>% rename(monthly_salary_28 = Var2,
                                                       company_size_23 = Var1)

levels(relation_ordinal_Q23_28$monthly_salary_28) <- gsub(" - ", "\n", levels(relation_ordinal_Q23_28$monthly_salary_28))




relation_ordinal_Q23_28 <- relation_ordinal_Q23_28 %>%
  # prepare text for tooltip
  mutate(text = paste(Freq, "respondent/s")) 



names(colours_6) <- levels(relation_ordinal_Q23_28$company_size_23)
colScale <- scale_colour_manual(name = "company_size_23",values = colours_6)

bubble_Q23_28_resp = ggplot(data = relation_ordinal_Q23_28, aes(x=company_size_23, y=monthly_salary_28, color=company_size_23)) +
  geom_point_interactive(aes(size=ifelse(Freq==0, NA, Freq),
                             tooltip = text)) +
  colScale +
  labs(x = "", y = "") +
  bubble_plotly_theme()



girafe(ggobj = bubble_Q23_28_resp, 
       fonts = list(sans = "Arial"),
       width_svg = 6, 
       height_svg = 4,
       options = list(
         opts_tooltip(use_fill = TRUE)))
```

:::

::: {.panel}

##### Only Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}
company_size_23_NA_remov <- Qs  %>%
  filter(profession_1 == "I am a student who is learning to code") %>%
  mutate(company_size_23_binned = recode(company_size_23, 
                                         "I don't know" = "NA",
                                         "I’m not sure"= "NA",  
                                         "more"= "NA",
                                         "N/A"= "NA"),
         monthly_salary_28_6 = recode(monthly_salary_28,
                                      "Greater than GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 20,000 - GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 15,000 - GHS 20,000"= "Greater than GHS 15,000",
                                      "GHS 10,000 - GHS 15,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 8,000 - GHS 10,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 6,000 - GHS 8,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 5,000 - GHS 6,000"= "GHS 3,500 - GHS 8,000",
                                      "GHS 4,000 - GHS 5,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 3,500 - GHS 4,000"= "GHS 3,500 - GHS 8,000")) %>%
  filter(company_size_23_binned != "NA") 

company_size_23_NA_remov$company_size_23_binned<- droplevels(company_size_23_NA_remov$company_size_23_binned)


runningcounts.df <- as.data.frame(table(company_size_23_NA_remov$company_size_23_binned, company_size_23_NA_remov$monthly_salary_28_6))

relation_ordinal_Q23_28 <- runningcounts.df %>% rename(monthly_salary_28 = Var2,
                                                       company_size_23 = Var1)

levels(relation_ordinal_Q23_28$monthly_salary_28) <- gsub(" - ", "\n", levels(relation_ordinal_Q23_28$monthly_salary_28))




relation_ordinal_Q23_28 <- relation_ordinal_Q23_28 %>%
  # prepare text for tooltip
  mutate(text = paste(Freq, "respondent/s")) 


# Heatmap as an alternative
# ggHeatmap(relation_ordinal_Q23_28,
#           aes(x = company_size_23,
#               y = monthly_salary_28,
#               fill = Freq),
#           addlabel=FALSE,
#           interactive=TRUE,
#           tooltip = text)


names(colours_6) <- levels(relation_ordinal_Q23_28$company_size_23)
colScale <- scale_colour_manual(name = "company_size_23",values = colours_6)

bubble_Q23_28_resp = ggplot(data = relation_ordinal_Q23_28, aes(x=company_size_23, y=monthly_salary_28, color=company_size_23)) +
  geom_point_interactive(aes(size=ifelse(Freq==0, NA, Freq),
                             tooltip = text)) +
  colScale +
  labs(x = "", y = "") +
  bubble_plotly_theme()



girafe(ggobj = bubble_Q23_28_resp, 
       fonts = list(sans = "Arial"),
       width_svg = 6, 
       height_svg = 4,
       options = list(
         opts_tooltip(use_fill = TRUE)))
```
:::

::::

#### Relationship between Highest Education and Salary of Professional Developers

A Bachelor's degree is associated with broad range of salary options. There are respondents With a Bachelors degree who receieve salaries as high as respondents with Master's degree. However, people with a Master's degree tend to have higher salaries than others on average.

```{r, fig.align='center', out.width = '100%', fig.height=3, warnings = FALSE}
# Variables: highest_edu_18, monthly_salary_28

monthly_salary_28_Prof <- Qs  %>%
  filter(profession_1 == "I am a developer by profession",
         highest_edu_18 == "Bachelor" | 
           highest_edu_18 == "Master" |
           highest_edu_18 == "Secondary High School" |
           highest_edu_18 == "Higher National Diploma") %>%
  mutate(monthly_salary_28_6 = recode(monthly_salary_28,
                                      "Greater than GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 20,000 - GHS 25,000" = "Greater than GHS 15,000",
                                      "GHS 15,000 - GHS 20,000"= "Greater than GHS 15,000",
                                      "GHS 10,000 - GHS 15,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 8,000 - GHS 10,000"= "GHS 8,000 - GHS 15,000",
                                      "GHS 6,000 - GHS 8,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 5,000 - GHS 6,000"= "GHS 3,500 - GHS 8,000",
                                      "GHS 4,000 - GHS 5,000"=  "GHS 3,500 - GHS 8,000",
                                      "GHS 3,500 - GHS 4,000"= "GHS 3,500 - GHS 8,000"))


monthly_salary_28_Prof$highest_edu_18<- droplevels(monthly_salary_28_Prof$highest_edu_18)
# levels(monthly_salary_28_Prof$highest_edu_18)

runningcounts.df <- as.data.frame(table(monthly_salary_28_Prof$highest_edu_18, monthly_salary_28_Prof$monthly_salary_28_6))

relation_ordinal_Q18_28 <- runningcounts.df %>% rename(monthly_salary_28 = Var2,
                                                       highest_edu_18 = Var1)

levels(relation_ordinal_Q18_28$monthly_salary_28) <- gsub(" - ", "\n", levels(relation_ordinal_Q18_28$monthly_salary_28))
levels(relation_ordinal_Q18_28$highest_edu_18) <- gsub(" ", "\n", levels(relation_ordinal_Q18_28$highest_edu_18))



relation_ordinal_Q18_28 <- relation_ordinal_Q18_28 %>%
  # prepare text for tooltip
  mutate(text = paste(Freq, "respondent/s")) 


names(colours_6) <- levels(relation_ordinal_Q18_28$highest_edu_18)
colScale <- scale_colour_manual(name = "highest_edu_18",values = colours_6)

bubble_Q18_28_resp = ggplot(data = relation_ordinal_Q18_28, aes(x=highest_edu_18, y=monthly_salary_28, color=highest_edu_18)) +
  geom_point_interactive(aes(size=ifelse(Freq==0, NA, Freq),
                             tooltip = text)) +
  colScale +
  labs(x = "", y = "") +
  bubble_plotly_theme()



girafe(ggobj = bubble_Q18_28_resp, 
       fonts = list(sans = "Arial"),
       width_svg = 6, 
       height_svg = 4,
       options = list(
         opts_tooltip(use_fill = TRUE)))

# ggplotly(bubble_Q18_28, tooltip="text", width = 800, height = 600)

```


### Which of the following best describes your current job-seeking status?

60% of the respondents are open for new job opportunities and one in three respondents is even actively looking for a job.

```{r, fig.align='center', out.width = '100%', fig.height=b3}

basic_barplot(my_df = compute_perc(Qs, job_status_29 ), 
              a_factor = job_status_29 , its_values = perc, labels = perc_label, 
              title = "")

```

### In general, what drives you to look for a new job?

The main reason for developers is more income whereas for students it is gaining more experience and skills improvement.

::::: {.panelset}

::: {.panel}

##### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=b6}

drivers_for_new_jobs_30_split<-Qs%>% select(drivers_for_new_jobs_30, profession_1) %>%
  separate_rows(drivers_for_new_jobs_30, sep = ";")


drivers_for_new_jobs_30_split<-drivers_for_new_jobs_30_split[!(is.na(drivers_for_new_jobs_30_split$drivers_for_new_jobs_30) | drivers_for_new_jobs_30_split$drivers_for_new_jobs_30==""), ]

drivers_for_new_jobs_30_split <-drivers_for_new_jobs_30_split %>% 
  mutate( drivers_for_new_jobs_30 = recode(drivers_for_new_jobs_30, 
                                           "A job i love and how challenging it is" = "Other",
                                           "A Fulltime / Stable Employment " = "Other",
                                           "Building Credibility and Connections(Relationships). As well making my skills practical." = "Other",
                                           "Funds to seed stuffs" = "Other",
                                           "I am focussed on running my business." = "Other",
                                           "I have a self motivated personality" = "Other",
                                           "to be in an environment where am challenged beyond my current capabilities" = "Other"
  ))

basic_barplot(my_df = compute_perc(drivers_for_new_jobs_30_split, drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```
``

:::

::: {.panel}

##### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=b6}

basic_barplot(my_df = compute_perc(drivers_for_new_jobs_30_split %>% filter(profession_1 == "I am a developer by profession"), drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```
:::

::: {.panel}

##### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=b6}

basic_barplot(my_df = compute_perc(drivers_for_new_jobs_30_split %>% filter(profession_1 == "I am a student who is learning to code"), drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```
:::

::::

### When job searching, how do you learn more about a company?

The most popular method to find out more about a company among the respondents is researching company media. There also seem no differences between developers and students.

```{r, fig.align='center', out.width = '100%', fig.height=3}

how_learn_about_company_31_split<-Qs%>% select(how_learn_about_company_31, profession_1) %>%
  separate_rows(how_learn_about_company_31, sep = ";")


how_learn_about_company_31_split<-how_learn_about_company_31_split[!(is.na(how_learn_about_company_31_split$how_learn_about_company_31) | how_learn_about_company_31_split$how_learn_about_company_31==""), ]


how_learn_about_company_31_split <-how_learn_about_company_31_split %>% 
  mutate(how_learn_about_company_31 = recode(how_learn_about_company_31, 
                                             "Company website" = "Read company media, such as employees blogs or company culture videos",
                                             "Google & their website" = "Read company media, such as employees blogs or company culture videos",
                                             "linkedin" = "Read company media, such as employees blogs or company culture videos",
                                             "Online (Linkedin" = "Read company media, such as employees blogs or company culture videos",
                                             "Websites " = "Read company media, such as employees blogs or company culture videos",
                                             "Personal assessment.. Occasionally going to the company to see how they cooperate with one another " = "Make personal contact to company",
                                             "Just going there frequently just access their interpersonal relationships with each other " = "Make personal contact with company"))

basic_barplot(my_df = compute_perc(how_learn_about_company_31_split, how_learn_about_company_31 ), 
              a_factor = how_learn_about_company_31 , its_values = perc, labels = perc_label, 
              title = "")


```



## Developer Communities

2 in 3 respondents are members of a developer community.

```{r, fig.align='center', out.width = '100%', fig.height=1}

basic_barplot(my_df = compute_perc(Qs, dev_community_member_32 ), 
              a_factor = dev_community_member_32 , its_values = perc, labels = perc_label, 
              title = "Are you a member of any other online/offline developer communities?")

```


The three largest communities to which respondents indicated membership were DevCongress (13.6%), Facebook Developer Circle (9.6%), and the Hacklab Foundation (7.4%).  

```{r, fig.align='center', out.width = '100%', fig.height=5.5}
# Warning! Check that all communities are in the plot. Depends on size (height) of figure and scale. Wordcloud excludes if names are too large (typically Facebook Developer Circle or DevCongress)
# I tried to pick a scale where the smallest were still legible.

# note: respondents can be members of several communities,
# we just count them all.
counted_communities <- Qs %>% select(ID, which_dev_community_33) %>%
  separate_rows(which_dev_community_33, sep = ";") %>% 
  drop_na() %>% 
  count(which_dev_community_33) %>% 
  arrange(-n) %>%
  mutate(perc_of_resp = n/number_of_respondents*100)

wordcloud(words = counted_communities$which_dev_community_33, freq = counted_communities$n, min.freq = 1,
          max.words = 122, random.order = TRUE, rot.per = 0, scale = c(2.2,0.5),
          colors = HL_colors)

```









