---
title: "HACKATHON 2021"
output: 
  distill::distill_article:
    toc: true     # adding table of contents
    toc_depth: 2
    toc_float: true
    theme: theme.css
---




```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE) # we do not want to display the code

# Environment:

library(xaringanExtra) # to be able to have tabs. To install: install 'devtools' package, then run: devtools::install_github("gadenbuie/xaringanExtra")
xaringanExtra::use_panelset() # allow the creation of tabs in Distill (see https://github.com/rstudio/distill/issues/11)

library(tidyverse) # includes dplyr, tidyr, ggplot2, rio, stringr, ...


```

```{r}

# Data:

Qs <- rio::import("data/clean/clean_all_Qs.rds")
skills <- rio::import("data/clean/skills_final.csv") %>% 
  select(-V1) %>%  # artifact due to a save as .csv to remove (at least on a Mac?)
  mutate_all(na_if,"") # the column 'level' contains blanks ("") instead of NAs -> transform them to NAs.
 
number_of_respondents = dim(Qs)[1] # number of respondents (272) used to compute percentages.
number_of_prof_dev = dim(Qs %>% filter(profession_1 == "I am a developer by profession"))[1]
number_of_students = dim(Qs %>% filter(profession_1 == "I am a student who is learning to code"))[1]


```

```{r}

# Themes and Functions:

HL_colors = c("#610b70","#88b101","#eb1c96","#e98403","#454545")
HL_font = "sans" # !!! Should change and use Gotham !!!
main_fontsize = 7 

barplot_theme <- function(){
  # the definition of a custom ggplot theme for our barplots.
  theme_minimal() + 
    theme(
      plot.title = element_text(family = HL_font, size = main_fontsize + 2, color = "black"),
      # plot.subtitle = element_text(family = HL_font, size = main_fontsize),
      axis.title = element_text(family = HL_font, size = main_fontsize, color = "black"),
      axis.text = element_text(family = HL_font, size = main_fontsize, color = "black"),
      # plot.caption = element_text(hjust = 0, family = my_font, 
      #                             size = main_fontsize, face= "italic"), #Default is hjust=1 / 0
      plot.title.position = "plot",  # left-align title
      # plot.caption.position =  "plot", 
      panel.grid.minor.x = element_blank(),  # remove useless minor grid
      panel.grid.major.y = element_blank(),  # remove useless major grid in bar charts
      panel.grid.major.x = element_blank(),  # remove useless major grid in bar charts
      axis.title.x = element_blank(), # we do not want an y axis for the barplot, we have labels already
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
      # plot.margin = margin(0.1, 0.1, 0.1, 0.1, unit = "cm")
    )
}

basic_barplot <- function(my_df, a_factor, its_values, labels, title){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = fct_reorder({{a_factor}}, {{its_values}}), 
                             y = {{its_values}})) + 
    geom_col(fill = HL_colors[1]) + 
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.06*max({{its_values}}))),
      size = 2
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}

compute_perc <- function(my_df, a_factor){
  # a small function to count a factor and compute the percentages
  # also return the percentages as char for labelling.
  # USE THIS FUNCTION IF THE SUM OF THE ANSWERS IS 272! (including NA) -> not working if several possible choices.
  my_df %>%
    count({{a_factor}}) %>%
    mutate(
      perc = n/sum(n),
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% return()
  
}



```




# Technology and Tech Culture

## Q7. How frequently do you learn a new language or framework?

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, new_tool_7), 
              a_factor = new_tool_7, its_values = perc, labels = perc_label, 
              title = "")

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), new_tool_7), 
              a_factor = new_tool_7, its_values = perc, labels = perc_label, 
              title = "")

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), new_tool_7), 
              a_factor = new_tool_7, its_values = perc, labels = perc_label, 
              title = "")

```

:::

:::::




## Q14. What is the primary operating system in which you work?

### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```

### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```

### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```



## Q15. In which operating system would you rather work?


### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```

### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```

### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```


## Q16. What level of influence do you, personally, have over new technology purchases at your organization?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, purch_influence_16), 
              a_factor = purch_influence_16, its_values = perc, labels = perc_label, 
              title = "Q16. What level of influence do you, personally, have over new technology purchases at your organization?")

```

## Q17. When buying a new tool or software, how do you discover and research available solutions? Select all that apply.

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, solution_research_17), 
              a_factor = solution_research_17, its_values = perc, labels = perc_label, 
              title = "Q17. When buying a new tool or software, how do you discover and research available solutions?")

```


# Education, Work and Career

## Q18. Which of the following best describes the highest level of formal education that you have completed?

### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, highest_edu_18), 
              a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
              title = "")

```

### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), highest_edu_18), 
              a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
              title = "")

```

### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), highest_edu_18), 
              a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
              title = "")

```



## Q19. What was your primary field of study?

### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, prim_study_19), 
              a_factor = prim_study_19, its_values = perc, labels = perc_label, 
              title = "")

```

### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), prim_study_19), 
              a_factor = prim_study_19, its_values = perc, labels = perc_label, 
              title = "")

```

### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), prim_study_19), 
              a_factor = prim_study_19, its_values = perc, labels = perc_label, 
              title = "")

```



## Q20. How important is a formal education, such as a university degree in computer science, to your career?

### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, edu_importance_20), 
              a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
              title = "")

```

### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), edu_importance_20), 
              a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
              title = "")

```

### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), edu_importance_20), 
              a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
              title = "")

```



## Q21. From Q18, if you could go back and change your educational path (but end up in the same career), what would you change?

```{r, fig.align='center', out.width = '100%', fig.height=3}

compute_perc(Qs, change_edu_21 )

```

## Q22. How satisfied are you with your current job? (If you work multiple jobs, answer for one you spend the most hours on.)

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, job_satisfaction_22 ), 
              a_factor = job_satisfaction_22 , its_values = perc, labels = perc_label, 
              title = "Q22. How satisfied are you with your current job?")

```

## Q23. Approximately how many people are employed by the company or organization you currently work for?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, company_size_23 ), 
              a_factor = company_size_23 , its_values = perc, labels = perc_label, 
              title = "Q23. Approximately how many people are employed by the company or organization you currently work for?")

```

## Q24. On average, how many hours per week do you work? Please enter a whole number in the box.

```{r, fig.align='center', out.width = '100%', fig.height=3}

compute_perc(Qs, weekly_work_hours_24)

```


## Q25. How often do you work overtime or beyond the formal time expectation of your job?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, overtime_work_25 ), 
              a_factor = overtime_work_25 , its_values = perc, labels = perc_label, 
              title = "Q25. How often do you work overtime or beyond the formal time expectation of your job?")

```

## Q26. Do you think your company has a good on-boarding process? (By on-boarding, we mean the structured process of getting you settled in to your new role at a company)

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, company_onboarding_process_26 ), 
              a_factor = company_onboarding_process_26 , its_values = perc, labels = perc_label, 
              title = "26. Do you think your company has a good on-boarding process?")

```

## Q27. How could on-boarding at your company be improved?

```{r, fig.align='center', out.width = '100%', fig.height=3}

compute_perc(Qs, improve_onboarding_27 )

```

## Q28. How much is your monthly salary in Ghana Cedis?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, monthly_salary_28 ), 
              a_factor = monthly_salary_28 , its_values = perc, labels = perc_label, 
              title = "Q28. How much is your monthly salary in Ghana Cedis?")

```

## Q29. Which of the following best describes your current job-seeking status?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, job_status_29 ), 
              a_factor = job_status_29 , its_values = perc, labels = perc_label, 
              title = "Q29. Which of the following best describes your current job-seeking status?")

```

## Q30. In general, what drives you to look for a new job? Select all that apply.

### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```

### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```

### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```



## Q31. When job searching, how do you learn more about a company? Select all that apply.

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, how_learn_about_company_31 ), 
              a_factor = how_learn_about_company_31 , its_values = perc, labels = perc_label, 
              title = "Q31. When job searching, how do you learn more about a company?")

```

## Q32. Are you a member of any other online/offline developer communities?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, dev_community_member_32 ), 
              a_factor = dev_community_member_32 , its_values = perc, labels = perc_label, 
              title = "Q32. Are you a member of any other online/offline developer communities?")

```

## Q33. Developer communities you are a member of

```{r, fig.align='center', out.width = '100%', fig.height=3}

compute_perc(Qs, which_dev_community_33 )


```


# Skills


## Which programming, scripting and markup languages have you done extensive development work in over the past year? (Q.8)

* _HTML/CSS_ is most popular, followed by _javascript_, _python_ and _SQL_.
* not same for dev and student! -> interesting


```{r}

languages = c(
  "C",			
  "C#",			
  "C++",		
  "HTML/CSS",			
  "Java",			
  "JavaScript",			
  'Kotlin',			
  "PHP",			
  "Python",			
  "Ruby",			
  "SQL",			
  "Swift",			
  "TypeScript"
)

# we add characteristics of the respondents to the skills, 
# so that we can differentiate their poopularity by occupation or gender:
skills_with_characteristics <- skills %>% left_join(select(Qs, ID, profession_1, gender_35), by = c("id" = "ID"))

compute_perc_popular_skills <- function(the_skills, skillset, number_of_resp){
  # USE THIS FUNCTION for the skills
  the_skills %>%
    filter(tool %in% skillset) %>% # select only the languages from all the tools
    filter(level == "Worked with in PAST year" | level == "Both") %>%
    count(tool) %>%
    mutate(
      perc = n/number_of_resp,
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% 
    arrange(-perc) %>%  
    return()
}


```


::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics, 
                                                  skillset = languages,
                                                  number_of_resp = number_of_respondents), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a developer by profession"), 
                                                  skillset = languages,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc_popular_skills(the_skills = skills_with_characteristics %>% 
                                                    filter(profession_1 == "I am a student who is learning to code"), 
                                                  skillset = languages,
                                                  number_of_resp = number_of_prof_dev), 
              a_factor = tool, its_values = perc, labels = perc_label, 
              title = "")

```

:::

:::::

# Demographic Data: Who were the respondents? 

## What describes you best?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, profession_1), 
              a_factor = profession_1, its_values = perc, labels = perc_label, 
              title = "")
```

## Do you code as a hobby? 

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::

## Employment
<<<<<<< HEAD

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, employment_3), 
=======
```{r}

#Function for barplots that does not reorder factor levels
barplot_ordered <- function(my_df, a_factor, its_values, labels, title){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = {{a_factor}}, {{its_values}}, 
                             y = {{its_values}})) + 
    geom_col(fill = HL_colors[1]) + 
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.06*max({{its_values}}))),
      size = 2
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}

employment_data <- Qs %>%
  filter(employment_3 != "I prefer not to say") %>%
  mutate(
    employment_3 = factor(employment_3, 
                          levels(employment_3)[c(4,2,3,6,5,1)])
      )
```


::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(employment_data, employment_3), 
>>>>>>> vivien
              a_factor = employment_3, its_values = perc, labels = perc_label, 
              title = "")
```

<<<<<<< HEAD
## Where do the developers come from? 

=======
:::

::: {.panel}

### Men only {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(employment_data %>% filter(gender_35 == "Man"), employment_3), 
              a_factor = employment_3, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Women only {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(employment_data %>% filter(gender_35 == "Woman"), employment_3), 
              a_factor = employment_3, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::



## Where do the developers come from? 

Zoom in to get a more detailed picture. 

Hover over the individual markers to see the professional status of the respondents. 

>>>>>>> vivien
```{r}
library(leaflet)
library(tmaptools)

data_city <- Qs %>%
  select(ID, city_5, profession_1) %>%
  filter(!city_5 %in% c("0", "Cincinnati, OH", "Pune, India", "kutunse, Ghana", "Awomaso, Ghana")) 

#Geocoding
data_city_geo <- data_city %>%
  mutate(
    geocode = geocode_OSM(data_city$city_5, as.sf = TRUE, keep.unfound = TRUE)
  )

#Map for the cities
violet_icon <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/pointhi/leaflet-color-markers/master/img/marker-icon-violet.png", 
  iconWidth = 24,
  iconHeight = 32) 

data_city_geo$profession_1 <- recode(data_city_geo$profession_1, 
                        "I am a developer by profession" = "Professional Developers", 
                        "I am not primarily a developer, but I write code sometimes as part of my work" = "Coding at Work",
                        "I am a student who is learning to code" = "Students", 
                        "I code primarily as a hobby" = "Hobby", 
                        "I used to be a developer by profession, but no longer am" = "Retired")



```

```{r, fig.align= 'center', out.width= '100%'}
leaflet(data = data_city_geo) %>%
  addTiles() %>% 
  addMarkers(~ geocode$lon, ~ geocode$lat, 
             clusterOptions = markerClusterOptions(), 
             icon = violet_icon, label = ~profession_1)
```

## How old are the respondents? 
```{r}
<<<<<<< HEAD
#Function for barplots that does not reorder factor levels
barplot_ordered <- function(my_df, a_factor, its_values, labels, title){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = {{a_factor}}, {{its_values}}, 
                             y = {{its_values}})) + 
    geom_col(fill = HL_colors[1]) + 
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.06*max({{its_values}}))),
      size = 2
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}
=======
>>>>>>> vivien

Qs$age_range_6 <- factor(Qs$age_range_6, levels(Qs$age_range_6)[c(6:1)])

```

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, age_range_6), 
              a_factor = age_range_6, its_values = perc, labels = perc_label, 
              title = "")

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}


barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), age_range_6), 
              a_factor = age_range_6, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), age_range_6), 
              a_factor = age_range_6, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::


## Gender 

### What Gender do you identify as?

<<<<<<< HEAD
Greater share of women among the students than the professional developers. 
=======
There is a  greater share of women among the students than among the professional developers. 
>>>>>>> vivien
Ghana's developer community may become more representative in the upcoming years. 

```{r}

#Removing NA

gender_data <- Qs %>%
<<<<<<< HEAD
  filter(gender_35 == "Woman" || gender_35 == "Man")
=======
  filter(gender_35 != "NA")
>>>>>>> vivien
```


::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(gender_data, gender_35), 
              a_factor = gender_35, its_values = perc, labels = perc_label, 
              title = "")

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}


barplot_ordered(my_df = compute_perc(gender_data %>% filter(profession_1 == "I am a developer by profession"), gender_35), 
              a_factor = gender_35, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(gender_data %>% filter(profession_1 == "I am a student who is learning to code"), gender_35), 
              a_factor = gender_35, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::


### Do you identify as transgender? 

<<<<<<< HEAD
```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, transgender_36), 
=======
```{r}
transgender_data <- Qs %>%
  filter(transgender_36 != "NA")
```


```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(transgender_data, transgender_36), 
>>>>>>> vivien
              a_factor = transgender_36, its_values = perc, labels = perc_label,
              title = "")

```


## Sexual orientation
<<<<<<< HEAD

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(Qs, sexual_orientation_37), 
=======
```{r}
sexual_orientation_data <- Qs %>%
    filter(sexual_orientation_37 != "NA" & sexual_orientation_37 != "Prefer not to say")

```

```{r, fig.align='center', out.width = '100%', fig.height=3}

barplot_ordered(my_df = compute_perc(sexual_orientation_data, sexual_orientation_37), 
>>>>>>> vivien
              a_factor = sexual_orientation_37, its_values = perc, labels = perc_label,
              title = "")

```

## Ethnicity 

```{r, fig.align='center', out.width = '100%', fig.height=3}

<<<<<<< HEAD
barplot_ordered(my_df = compute_perc(Qs, ethnicity_38), 
=======
barplot_ordered(my_df = compute_perc(Qs %>% filter(ethnicity_38 != "NA" & ethnicity_38 != "I prefer not to answer"), ethnicity_38), 
>>>>>>> vivien
              a_factor = ethnicity_38, its_values = perc, labels = perc_label,
              title = "")

```

## Do you have any dependents you care for? 
<<<<<<< HEAD
=======
```{r}
dependents_data <- Qs %>%
  filter(dependents_41 != "I prefer not to say" & dependents_41 != "NA")
```
>>>>>>> vivien

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

<<<<<<< HEAD
barplot_ordered(my_df = compute_perc(Qs, dependents_41), 
=======
barplot_ordered(my_df = compute_perc(dependents_data, dependents_41), 
>>>>>>> vivien
              a_factor = dependents_41, its_values = perc, labels = perc_label, 
              title = "")

```


:::

::: {.panel}

### Only Men {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}


<<<<<<< HEAD
barplot_ordered(my_df = compute_perc(Qs %>% filter(gender_35 == "Man"), dependents_41), 
=======
barplot_ordered(my_df = compute_perc(dependents_data %>% filter(gender_35 == "Man"), dependents_41), 
>>>>>>> vivien
              a_factor = dependents_41, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Only Women {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

<<<<<<< HEAD
barplot_ordered(my_df = compute_perc(Qs %>% filter(gender_35 == "Woman"), dependents_41), 
=======
barplot_ordered(my_df = compute_perc(dependents_data %>% filter(gender_35 == "Woman"), dependents_41), 
>>>>>>> vivien
              a_factor = dependents_41, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::





