---
title: "Hacklab Developer Census 2020"
description: "Preliminary Draft - `r Sys.Date()`"
output: 
  html_document:  
    # instead of distill::distill_article:
    toc: true
    toc_float: true # not working in distill...
    toc_depth: 2
    toc_collapsed: true
    # theme: paper
    css: css_theme/css_from_scratch.css
    # css: css_theme/Splendor.css
    # css: css_theme/readthedown/readthedown.css
    # css: css_theme/stackoverflow.css # not working...
---


```{css toc-content, echo = FALSE}
/* add some space before the TOC */
/* for no apparent reason, this is working here but not in the .css... */
#TOC {
  margin-top: 5rem;
}
```



```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE) # we do not want to display the code

# Environment:

library(xaringanExtra) # to be able to have tabs. To install: install 'devtools' package, then run: devtools::install_github("gadenbuie/xaringanExtra")
xaringanExtra::use_panelset() # allow the creation of tabs in Distill (see https://github.com/rstudio/distill/issues/11)

library(tidyverse) # includes dplyr, tidyr, ggplot2, rio, stringr, ...

library(leaflet) # for the interactive map
library(tmaptools) # for geocoding

library(wordcloud) # for the wordcloud at Q33

library(tidyr) # for Q30 and Q31

library(ggiraph) # for "interactive" plots

library(ggchicklet) # to round edges of barplots. install.packages("ggchicklet", repos = "https://cinc.rud.is")

```

```{r}

# Data:

Qs <- rio::import("../../data/clean/clean_all_Qs.rds") %>%
  # reodering certain factors for the plots: (use barplot_ordered() for plotting)
  mutate(
    purch_influence_16 = fct_relevel(purch_influence_16,
                                     "I have little or no influence",
                                     "I have some influence",
                                     "I have a great deal of influence"
    ),    
    job_satisfaction_22 = fct_relevel(job_satisfaction_22,
                                      "Very dissatisfied",
                                      "Slightly dissatisfied",
                                      "Neither satisfied nor dissatisfied",
                                      "Slightly satisfied",
                                      "Very satisfied"),
    edu_importance_20 = fct_relevel(edu_importance_20,
                                    "Not at all important/not necessary",
                                    "Somewhat important",
                                    "Fairly important",
                                    "Very important",
                                    "Critically important"),
    company_size_23 = fct_relevel(company_size_23,
                                  "Over 500",
                                  "Over 100",
                                  "Below 100",
                                  "Below 20",
                                  "Below 10",
                                  "1"),
    overtime_work_25 = fct_relevel(overtime_work_25,
                                   "Often: 1-2 days per week",
                                   "Frequently: 3 or more days per week",
                                   "Sometimes: 1-2 days per month but less than weekly",
                                   "Occasionally: 1-2 days per quarter but less than monthly",
                                   "Rarely: 1-2 days per year or less",
                                   "Never"),
    monthly_salary_28 = fct_relevel(monthly_salary_28,
                                    "Greater than GHS 25,000",
                                    "GHS 20,000 - GHS 25,000",
                                    "GHS 15,000 - GHS 20,000",
                                    "GHS 10,000 - GHS 15,000",
                                    "GHS 8,000 - GHS 10,000",
                                    "GHS 6,000 - GHS 8,000",
                                    "GHS 5,000 - GHS 6,000",
                                    "GHS 4,000 - GHS 5,000",
                                    "GHS 3,500 - GHS 4,000",
                                    "GHS 3,000 - GHS 3,500",
                                    "GHS 2,500 - GHS 3,000",
                                    "GHS 2,000 - GHS 2,500",
                                    "GHS 1,500 - GHS 2,000",
                                    "Less than GHS 1,500"),
    prim_study_19 = recode(prim_study_19, 
                           "Computer_Science_Engineering" = "Computer Science/Engineering",
                           "Information_Technology" = "Information Technology",
                           "Other_Engineering" = "Other Engineering",
                           "Health_Science" = "Health Science",
                           "Social_Science" = "Social Science",
                           "Web_Development_Web_Design" = "Web-Development/Web-Design",
                           "General_Science" = "General Science",
                           "Electrical_Engineering" = "Electrical Engineering",
                           "Visual_Arts" = "Visual Arts",
                           "Agricultural_Science" = "Agricultural Science",
                           "Graphic_Design" = "Graphic Design",
                           "General_Arts" = "General Arts",
                           "Mathematics_Statistics" = "Mathematics/Statistics"),
    prim_study_19 = fct_relevel(prim_study_19,
                                "Education",
                                "Social Science",
                                "Graphic Design",
                                "Architecture",
                                "Visual Arts",
                                "Business",
                                "General Arts",
                                "General Science",
                                "Health Science",
                                "Chemistry",
                                "Other Engineering",
                                "Physics",
                                "Agricultural Science",
                                "Web-Development/Web-Design",
                                "Mathematics/Statistics",
                                "Electrical Engineering",
                                "Information Technology",
                                "Computer Science/Engineering"),   
    highest_edu_18 = recode(highest_edu_18, 
                            "Secondary_High_School" = "Secondary High School",
                            "Basic_Education" = "Basic Education",
                            "Higher_National_Diploma" = "Higher National Diploma",
                            "Professional_Diploma" = "Professional Diploma",
                            "Teacher_Diploma" = "Teacher Diploma",
                            "College" = "Bachelor"),
    highest_edu_18 = fct_relevel(highest_edu_18,
                                 "Basic Education",
                                 "Secondary High School",
                                 "Teacher Diploma",
                                 "Professional Diploma",
                                 "Higher National Diploma",
                                 "Bachelor",
                                 "Master")
  )

skills <- rio::import("../../data/clean/skills_final.csv") %>% 
  select(-V1) %>%  # artifact due to a save as .csv to remove (at least on a Mac?)
  mutate_all(na_if,"") # the column 'level' contains blanks ("") instead of NAs -> transform them to NAs.
# we add characteristics of the respondents to the skills, 
# so that we can differentiate their popularity by occupation or gender:
skills_with_characteristics <- skills %>% left_join(select(Qs, ID, profession_1, gender_35), by = c("id" = "ID"))

number_of_respondents = dim(Qs)[1] # number of respondents (272) used to compute percentages.
number_of_prof_dev = dim(Qs %>% filter(profession_1 == "I am a developer by profession"))[1]
number_of_students = dim(Qs %>% filter(profession_1 == "I am a student who is learning to code"))[1]


```

```{r}

# Themes and Functions:

highlight_col = "#d4145a"
text_col = "#777777"
header_col = "#222222"
font_in_viz = "sans" # !!! Should change and use Gotham !!!
text_col_in_viz = "black"
main_fontsize = 10
HL_colors = c(text_col, "#610b70","#88b101","#eb1c96","#e98403","#454545")
geom_text_size = 3.9


barplot_theme <- function(){
  # the definition of a custom ggplot theme for our barplots.
  theme_minimal() + 
    theme(
      plot.title = element_text(family = font_in_viz, size = main_fontsize + 2, color = text_col_in_viz),
      # plot.subtitle = element_text(family = font_in_viz, size = main_fontsize),
      axis.title = element_text(family = font_in_viz, size = main_fontsize, color = text_col_in_viz),
      axis.text = element_text(family = font_in_viz, size = main_fontsize, color = text_col_in_viz),
      # plot.caption = element_text(hjust = 0, family = my_font, 
      #                             size = main_fontsize, face= "italic"), #Default is hjust=1 / 0
      plot.title.position = "plot",  # left-align title
      # plot.caption.position =  "plot", 
      panel.grid.minor.x = element_blank(),  # remove useless minor grid
      panel.grid.major.y = element_blank(),  # remove useless major grid in bar charts
      panel.grid.major.x = element_blank(),  # remove useless major grid in bar charts
      axis.title.x = element_blank(), # we do not want an y axis for the barplot, we have labels already
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
      # plot.margin = margin(0.1, 0.1, 0.1, 0.1, unit = "cm")
    )
}

basic_barplot <- function(my_df, a_factor, its_values, labels, title){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = fct_reorder({{a_factor}}, {{its_values}}), 
                             y = {{its_values}})) + 
    geom_chicklet(radius = grid::unit(1, 'mm'), fill = HL_colors[1]) +  # NNNNNEEEEEEEEEWWWWW !!!!!!
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.06*max({{its_values}}))),
      size = geom_text_size
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}

basic_barplot_dark <- function(my_df, a_factor, its_values, labels, title){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = fct_reorder({{a_factor}}, {{its_values}}), 
                             y = {{its_values}})) + 
    geom_chicklet(radius = grid::unit(1, 'mm'), fill = header_col) +  # NNNNNEEEEEEEEEWWWWW !!!!!!
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.06*max({{its_values}}))),
      size = geom_text_size
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}

basic_barplot_pink <- function(my_df, a_factor, its_values, labels, title){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = fct_reorder({{a_factor}}, {{its_values}}), 
                             y = {{its_values}})) + 
    geom_chicklet(radius = grid::unit(1, 'mm'), fill = highlight_col) +  # NNNNNEEEEEEEEEWWWWW !!!!!!
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.06*max({{its_values}}))),
      size = geom_text_size
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}

#Function for barplot that does not override order
barplot_ordered <- function(my_df, a_factor, its_values, labels, title, tooltip, data_id){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = {{a_factor}},
                             y = {{its_values}},
                             tooltip = {{tooltip}}, 
                             data_id = {{data_id}})) + 
    geom_col_interactive(fill = HL_colors[1]) + 
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.06*max({{its_values}}))),
      size = geom_text_size
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}

compute_perc <- function(my_df, a_factor){
  # a small function to count a factor and compute the percentages
  # also return the percentages as char for labelling.
  # USE THIS FUNCTION IF THE SUM OF THE ANSWERS IS 272! (including NAs) -> not working if several possible choices.
  my_df %>%
    count({{a_factor}}) %>%
    drop_na() %>% # always in percentage of the respondents to the question.
    mutate(
      perc = n/sum(n),
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% return()
  
}

compute_perc_popular_skills <- function(the_skills, skillset, number_of_resp){
  # USE THIS FUNCTION for the popular skills (last year + both)
  # NOTE: we use all respondents, not only those who responded. -> some respondents did not enter any skills. -> not optimal, but annoying to correct?
  the_skills %>%
    filter(tool %in% skillset) %>% # select only the languages from all the tools
    filter(level == "Worked with in PAST year" | level == "Both") %>%
    count(tool) %>%
    mutate(
      perc = n/number_of_resp, 
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% 
    arrange(-perc) %>%  
    return()
}

compute_perc_future_skills <- function(the_skills, skillset, number_of_resp){
  # USE THIS FUNCTION for the future, nexy year, skills (next year + both)
  # NOTE: we use all respondents, not only those who responded. -> some respondents did not enter any skills. -> not optimal, but annoying to correct?
  the_skills %>%
    filter(tool %in% skillset) %>% # select only the languages from all the tools
    filter(level == "Want to work with NEXT year" | level == "Both") %>%
    count(tool) %>%
    mutate(
      perc = n/number_of_resp, 
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% 
    arrange(-perc) %>%  
    return()
}

fig_height_factor = 2


```


# Basic Information {.header1} 

## What describes you best?

The two biggest subgroups among the respondents are professional developers as well as students. Additionally, there are respondents coding as a part of their work, coding as a hobby, as well as former developers. 

```{r}
# THIS IS A TEST: TRY TO SEE IF WE CAN DYNAMICALLY DEFINE FIGURE SIZES!
next_fig_height = dim(compute_perc(Qs, profession_1))[1] / fig_height_factor 
```

```{r, fig.align='center', out.width = '100%', fig.height=next_fig_height}

basic_barplot(my_df = compute_perc(Qs, profession_1), 
              a_factor = profession_1, its_values = perc, labels = perc_label, 
              title = "")
```

## Do you code as a hobby? 

Most of the respondents code as a hobby. Interestingly, professional developers seem to code for a hobby less often than students which may be related to their reduced time availability. 

```{r}
# THIS IS A TEST: TRY TO SEE IF WE CAN DYNAMICALLY DEFINE FIGURE SIZES!
next_fig_height = dim(compute_perc(Qs, hobby_coding_2))[1] / fig_height_factor
```


::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=next_fig_height}

basic_barplot(my_df = compute_perc(Qs, hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=next_fig_height}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=next_fig_height}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), hobby_coding_2), 
              a_factor = hobby_coding_2, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::

## Employment

```{r}

employment_data <- Qs %>%
  filter(employment_3 != "I prefer not to say") %>%
  mutate(
    employment_3 = factor(employment_3, 
                          levels(employment_3)[c(4,2,3,6,5,1)])
  )

# THIS IS A TEST: TRY TO SEE IF WE CAN DYNAMICALLY DEFINE FIGURE SIZES!
next_fig_height = dim(compute_perc(employment_data, employment_3))[1] / fig_height_factor 

```

Most of the respondents are employed full-time. Furthermore, there is a great share of students. Sadly, there is also a significant share of unemployed developers looking for work. 

When it comes to gender differences, there is a greater share of students among the female respondents. In addition, women seem to be self-employed less often when compared with men. 

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=next_fig_height}

basic_barplot_dark(my_df = compute_perc(employment_data, employment_3), 
              a_factor = employment_3, its_values = perc, labels = perc_label,
              title = "")

```


:::

::: {.panel}

### Men only {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=next_fig_height}

basic_barplot_dark(my_df = compute_perc(employment_data %>% filter(gender_35 == "Man"), employment_3), 
                a_factor = employment_3, its_values = perc, labels = perc_label,
                title = "")

```

:::

::: {.panel}

### Women only {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=next_fig_height}

basic_barplot_dark(my_df = compute_perc(employment_data %>% filter(gender_35 == "Woman"), employment_3), 
                a_factor = employment_3, its_values = perc, labels = perc_label,
                title = "")

```

:::

::::



## Where do the developers come from? 

### Region

Most respondents come from the greater Accra region. However, the extent of concentration in Accra seems to be larger for professional developers than for students. 

```{r}

region_data <- Qs %>%
  filter(region_4 != "Not in Ghana")

region_data$region_4 <- recode(region_data$region_4,
                               "Eastern Region, Ghana" = "Eastern Region", 
                               "Northern Region, Ghana" = "Northern Region", 
                               "Central Region, Ghana" = "Central Region", 
                               "Western Region, Ghana"= "Western Region"
)

# THIS IS A TEST: TRY TO SEE IF WE CAN DYNAMICALLY DEFINE FIGURE SIZES!
next_fig_height = dim(compute_perc(region_data, region_4))[1] / fig_height_factor

```


::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=next_fig_height}


basic_barplot_pink(my_df = compute_perc(region_data, region_4), 
              a_factor = region_4, its_values = perc, labels = perc_label,
              title = "")

```


:::

::: {.panel}

### Professional Developers {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=next_fig_height}

basic_barplot_pink(my_df = compute_perc(region_data %>% filter(profession_1 == "I am a developer by profession"), region_4), 
              a_factor = region_4, its_values = perc, labels = perc_label,
              title = "")

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=next_fig_height}

basic_barplot_pink(my_df = compute_perc(region_data %>% filter(profession_1 == "I am a student who is learning to code"), region_4), 
              a_factor = region_4, its_values = perc, labels = perc_label,
              title = "")

```

:::

::::




### City

Zoom in to get a more detailed picture. 

Hover over the individual markers to see the professional status of the respondents. 

```{r, fig.align= 'center', out.width= '100%', fig.height=6.5}

# the geocoding was done separately in data_cleaning/Geocoding.R
# (to not lose time everytime we knitr)
# we load the result:
data_city_geo <- rio::import("../../data/clean/data_city_geo.rds") 

#Map for the cities
violet_icon <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/pointhi/leaflet-color-markers/master/img/marker-icon-violet.png", 
  iconWidth = 24,
  iconHeight = 32) 

leaflet(data = data_city_geo) %>%
  addTiles() %>% 
  addMarkers(~ geocode$lon, ~ geocode$lat, 
             clusterOptions = markerClusterOptions(), 
             icon = violet_icon, label = ~profession_1)
```

## How old are the respondents? 

Most respondents are between 20 and 30. 

While the students are very young in most cases, the professional developers display a little more variance in their ages. However, we have a very young sample at hand.

```{r}
Qs$age_range_6 <- factor(Qs$age_range_6, levels(Qs$age_range_6)[c(6:1)])

```

::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}

p <- barplot_ordered(my_df = compute_perc(Qs, age_range_6), 
                a_factor = age_range_6, its_values = perc, labels = perc_label, 
                tooltip = perc_label, data_id = perc,
                title = "") + scale_x_discrete(limits=rev)
girafe(ggobj = p,
       options = list(
         opts_sizing(rescale = FALSE), # else, rescale differently...
         opts_hover_inv(css = "opacity:0.1;"),
         opts_hover(css = "fill:#d4145a;")
       )
)

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=3}


p <- barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), age_range_6), 
                a_factor = age_range_6, its_values = perc, labels = perc_label, 
                tooltip = perc_label, data_id = perc,
                title = "") + scale_x_discrete(limits=rev)
girafe(ggobj = p,
       options = list(
         opts_sizing(rescale = FALSE), # else, rescale differently...
         opts_hover_inv(css = "opacity:0.1;"),
         opts_hover(css = "fill:#d4145a;")
       )
)

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=3}


p <- barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), age_range_6), 
                a_factor = age_range_6, its_values = perc, labels = perc_label, 
                tooltip = perc_label, data_id = perc,
                title = "") + scale_x_discrete(limits=rev)
girafe(ggobj = p,
       options = list(
         opts_sizing(rescale = FALSE), # else, rescale differently...
         opts_hover_inv(css = "opacity:0.1;"),
         opts_hover(css = "fill:#d4145a;")
       )
)

```

:::

::::







# Technology and Tech Culture {.header1} 

## How frequently do you learn a new language or framework?



::::: {.panelset}

::: {.panel}

### All Respondents {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=2}

p <- barplot_ordered(my_df = compute_perc(Qs, new_tool_7), 
                a_factor = new_tool_7, its_values = perc, labels = perc_label, 
                tooltip = perc_label, data_id = perc,
                title = "") + scale_x_discrete(limits=rev)
girafe(ggobj = p,
       options = list(
         opts_sizing(rescale = FALSE), # else, rescale differently...
         opts_hover_inv(css = "opacity:0.1;"),
         opts_hover(css = "fill:#d4145a;")
       )
)

```


:::

::: {.panel}

### Professional Developers {.panel-name}
```{r, fig.align='center', out.width = '100%', fig.height=2}

p <- barplot_ordered(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), new_tool_7), 
                a_factor = new_tool_7, its_values = perc, labels = perc_label, 
                tooltip = perc_label, data_id = perc,
                title = "") + scale_x_discrete(limits=rev)
girafe(ggobj = p,
       options = list(
         # opts_hover_inv(css = "opacity:0.1;"),
         opts_hover(css = "fill:#d4145a;")
       )
)

```

:::

::: {.panel}

### Students {.panel-name}

```{r, fig.align='center', out.width = '100%', fig.height=2}

p <- barplot_ordered(my_df = compute_perc(Qs %>% 
                                            filter(profession_1 == "I am a student who is learning to code"), new_tool_7),
                a_factor = new_tool_7, its_values = perc, labels = perc_label, 
                tooltip = perc_label, data_id = perc,
                title = "") + scale_x_discrete(limits=rev)
girafe(ggobj = p,
       options = list(
         opts_hover_inv(css = "opacity:0.1;"),
         opts_hover(css = "fill:#d4145a;")
       )
)

```

:::

:::::






