---
title: "Technology and Tech Culture"
description: |
  Exploring and trying to find insights for questions 7 to 17
# author:
#   - name: Nora Jones 
#     url: https://example.com/norajones
#     affiliation: Spacely Sprockets
#     affiliation_url: https://example.com/spacelysprokets

date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(ggiraph)

Qs <- rio::import("../../data/clean/clean_all_Qs.rds")
skills <- rio::import("../../data/clean/skills_final.csv") %>% select(-V1)


# a theme for the plot
HL_colors = c("#610b70","#88b101","#eb1c96","#e98403","#454545")
HL_font = "Open Sans" # I dunno why I can't make Gotham work??? "Gotham Book"
main_fontsize = 7 # must be large if used primarily on slides 

barplot_theme <- function(){
  theme_minimal() + 
    theme(
      plot.title = element_text(family = HL_font, size = main_fontsize + 2, color = "black"),
      # plot.subtitle = element_text(family = HL_font, size = main_fontsize),
      axis.title = element_text(family = HL_font, size = main_fontsize, color = "black"),
      axis.text = element_text(family = HL_font, size = main_fontsize, color = "black"),
      # plot.caption = element_text(hjust = 0, family = my_font, 
      #                             size = main_fontsize, face= "italic"), #Default is hjust=1 / 0
      plot.title.position = "plot",  # left-align title
      # plot.caption.position =  "plot", 
      panel.grid.minor.x = element_blank(),  # remove useless minor grid
      panel.grid.major.y = element_blank(),  # remove useless major grid in bar charts
      panel.grid.major.x = element_blank(),  # remove useless major grid in bar charts
      axis.title.x = element_blank(), # we do not want an y axis for the barplot, we have labels already
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
      # plot.margin = margin(0.1, 0.1, 0.1, 0.1, unit = "cm")
    )
}



```


# Open Questions

* Should we try to compare the respondents to StackOverflow respondents? In other words, highlight the particularities of the Ghanaian tech community?
* Detail results for "All respondents", "Professional Developers", and eventually "Students"?
* Always plot in percentage and do not care about NAs? 




# Some Key Observations

* Q1 - respondents are mostly developers or students. Some also code as a hobby, or part-time in their job. (Nevertheless, most indicate coding as a hobby too in Q.2)


# A more detailed look

There are enough professional developers and students so that we could contrast them.  

```{r, fig.align='center', out.width = '100%', fig.height=3}

# need to build a plotting function for the bar chart. 

current_plot <- Qs %>% 
  count(profession_1) %>%
  mutate(
    perc = n/sum(n),
    perc_label = paste0(as.character(round(perc*100,1)), "%")
  )%>%
  ggplot(aes(x = fct_reorder(profession_1, perc), y = perc, label = perc_label)) + 
  geom_col(fill = HL_colors[1]) + 
  geom_text(
    aes(label = perc_label, y = perc + (0.06*max(perc))),
    size = 2
  ) +
  scale_y_continuous(labels = scales::percent) +
  barplot_theme() + 
  coord_flip() +
  labs(title = "Q.1 Which of the following options best describes you today?") +
  xlab("") +
  ylab("")

current_plot


```



### Q7. How frequently do you learn a new language or framework?

```{r}

Qs %>% count(new_tool_7)


```










