---
title: "Technology and Tech Culture + Education, Work, Career"
description: "Exploring and finding insights for questions 7 to 17 + 18 to 33"
# author:
#   - name: Nora Jones 
#     url: https://example.com/norajones
#     affiliation: Spacely Sprockets
#     affiliation_url: https://example.com/spacelysprokets

date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    theme: test_custom_theme.css # loading a custom css (saved in same folder)
    toc: true # table of contents
    toc_depth: 2 # only include titles level # and ## in toc

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(ggiraph) # to try "interactive" plots

# library(xaringanExtra)
# xaringanExtra::use_panelset() # https://github.com/rstudio/distill/issues/11
# ==> tabset is not working in Distill articles, see: https://github.com/rstudio/distill/issues/11
# there would be a workaround with xaringanextra: https://pkg.garrickadenbuie.com/xaringanExtra/#/panelset?id=use-in-r-markdown
# not working with my R version, but that might be a good argument to not use Distill!

Qs <- rio::import("../../data/clean/clean_all_Qs.rds")
skills <- rio::import("../../data/clean/skills_final.csv") %>% select(-V1)

# a theme for the plot
HL_colors = c("#610b70","#88b101","#eb1c96","#e98403","#454545")
HL_font = "sans" # !!! Should change and use Gotham !!!
main_fontsize = 7 

barplot_theme <- function(){
  # the definition of a custom ggplot theme for our barplots.
  theme_minimal() + 
    theme(
      plot.title = element_text(family = HL_font, size = main_fontsize + 2, color = "black"),
      # plot.subtitle = element_text(family = HL_font, size = main_fontsize),
      axis.title = element_text(family = HL_font, size = main_fontsize, color = "black"),
      axis.text = element_text(family = HL_font, size = main_fontsize, color = "black"),
      # plot.caption = element_text(hjust = 0, family = my_font, 
      #                             size = main_fontsize, face= "italic"), #Default is hjust=1 / 0
      plot.title.position = "plot",  # left-align title
      # plot.caption.position =  "plot", 
      panel.grid.minor.x = element_blank(),  # remove useless minor grid
      panel.grid.major.y = element_blank(),  # remove useless major grid in bar charts
      panel.grid.major.x = element_blank(),  # remove useless major grid in bar charts
      axis.title.x = element_blank(), # we do not want an y axis for the barplot, we have labels already
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
      # plot.margin = margin(0.1, 0.1, 0.1, 0.1, unit = "cm")
    )
}

basic_barplot <- function(my_df, a_factor, its_values, labels, title){
  # a function to plot a standard barplot
  # uses the barplot_theme defined above
  # needs an (ordered) factor as input, and its values
  # returns a ggplot
  current_plot <- ggplot(data = my_df, 
                         aes(x = fct_reorder({{a_factor}}, {{its_values}}), 
                             y = {{its_values}})) + 
    geom_col(fill = HL_colors[1]) + 
    geom_text(
      aes(label = {{labels}}, y = {{its_values}} + (0.06*max({{its_values}}))),
      size = 2
    ) +
    scale_y_continuous(labels = scales::percent) +
    barplot_theme() + 
    coord_flip() +
    labs(title = stringr::str_wrap(title,100)) + # automated text wrap in title
    xlab("") +
    ylab("")
  return(current_plot)
}

compute_perc <- function(my_df, a_factor){
  # a small function to count a factor and compute the percentages
  # also return the percentages as char for labelling.
  my_df %>%
    count({{a_factor}}) %>%
    mutate(
      perc = n/sum(n),
      perc_label = paste0(as.character(round(perc*100,1)), "%")
    ) %>% return()
  
}

```


# TODO
* skills (Q8-13)
* clean and group Q17, Q21, Q27, Q30, Q31, Q33 (wordcloud again?)
* histogram for Q24
* develop some questions by gender
* the barplot function should have an option to reorder the factors or not (e.g. Q22, Q23)
* the barplots should have the same y-axis, at least for the same tabs.
* change font to Gotham
* try and make ggiraph work
* check what StackOverflow did
* idea: 2D plot with Salary VS Education, Salary VS Field of study, etc.  


# Open Questions

* Should we try to compare the respondents to StackOverflow respondents? In other words, highlight the particularities of the Ghanaian tech community?

# Solved Questions
* Detail results for "All respondents", "Professional Developers", and eventually "Students"? ==> YES
* Always plot in percentage and do not care about NAs? ==> YES. __[NOTE]__: for now, the NA are still included in the plots to not lose this info when analyzing. For the final plots, they will have to be removed.  


# Some Key Observations

* Q1 - respondents are mostly developers or students. Some also code as a hobby, or part-time in their job. (Nevertheless, most indicate coding as a hobby too in Q.2)


# A more detailed look

There are enough professional developers and students so that we could contrast them.  

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, profession_1), 
              a_factor = profession_1, its_values = perc, labels = perc_label, 
              title = "Q.1 Which of the following options best describes you today?")

```

## Technology and Tech Culture

### Q7. How frequently do you learn a new language or framework?


#### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, new_tool_7), 
              a_factor = new_tool_7, its_values = perc, labels = perc_label, 
              title = "")

```

#### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), new_tool_7), 
              a_factor = new_tool_7, its_values = perc, labels = perc_label, 
              title = "")

```

#### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), new_tool_7), 
              a_factor = new_tool_7, its_values = perc, labels = perc_label, 
              title = "")

```


### Q14. What is the primary operating system in which you work?

#### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```

#### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```

#### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), prim_opsyst_14), 
              a_factor = prim_opsyst_14, its_values = perc, labels = perc_label, 
              title = "")

```



### Q15. In which operating system would you rather work?


#### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```

#### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```

#### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), rather_opsyst_15), 
              a_factor = rather_opsyst_15, its_values = perc, labels = perc_label, 
              title = "")

```


### Q16. What level of influence do you, personally, have over new technology purchases at your organization?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, purch_influence_16), 
              a_factor = purch_influence_16, its_values = perc, labels = perc_label, 
              title = "Q16. What level of influence do you, personally, have over new technology purchases at your organization?")

```

### Q17. When buying a new tool or software, how do you discover and research available solutions? Select all that apply.

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, solution_research_17), 
              a_factor = solution_research_17, its_values = perc, labels = perc_label, 
              title = "Q17. When buying a new tool or software, how do you discover and research available solutions?")

```


## Education, Work and Career

### Q18. Which of the following best describes the highest level of formal education that you have completed?

#### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, highest_edu_18), 
              a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
              title = "")

```

#### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), highest_edu_18), 
              a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
              title = "")

```

#### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), highest_edu_18), 
              a_factor = highest_edu_18, its_values = perc, labels = perc_label, 
              title = "")

```



### Q19. What was your primary field of study?

#### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, prim_study_19), 
              a_factor = prim_study_19, its_values = perc, labels = perc_label, 
              title = "")

```

#### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), prim_study_19), 
              a_factor = prim_study_19, its_values = perc, labels = perc_label, 
              title = "")

```

#### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), prim_study_19), 
              a_factor = prim_study_19, its_values = perc, labels = perc_label, 
              title = "")

```



### Q20. How important is a formal education, such as a university degree in computer science, to your career?

#### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, edu_importance_20), 
              a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
              title = "")

```

#### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), edu_importance_20), 
              a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
              title = "")

```

#### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), edu_importance_20), 
              a_factor = edu_importance_20, its_values = perc, labels = perc_label, 
              title = "")

```



### Q21. From Q18, if you could go back and change your educational path (but end up in the same career), what would you change?

```{r, fig.align='center', out.width = '100%', fig.height=3}

compute_perc(Qs, change_edu_21 )

```

### Q22. How satisfied are you with your current job? (If you work multiple jobs, answer for one you spend the most hours on.)

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, job_satisfaction_22 ), 
              a_factor = job_satisfaction_22 , its_values = perc, labels = perc_label, 
              title = "Q22. How satisfied are you with your current job?")

```

### Q23. Approximately how many people are employed by the company or organization you currently work for?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, company_size_23 ), 
              a_factor = company_size_23 , its_values = perc, labels = perc_label, 
              title = "Q23. Approximately how many people are employed by the company or organization you currently work for?")

```

### Q24. On average, how many hours per week do you work? Please enter a whole number in the box.

```{r, fig.align='center', out.width = '100%', fig.height=3}

compute_perc(Qs, weekly_work_hours_24)

```


### Q25. How often do you work overtime or beyond the formal time expectation of your job?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, overtime_work_25 ), 
              a_factor = overtime_work_25 , its_values = perc, labels = perc_label, 
              title = "Q25. How often do you work overtime or beyond the formal time expectation of your job?")

```

### Q26. Do you think your company has a good on-boarding process? (By on-boarding, we mean the structured process of getting you settled in to your new role at a company)

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, company_onboarding_process_26 ), 
              a_factor = company_onboarding_process_26 , its_values = perc, labels = perc_label, 
              title = "26. Do you think your company has a good on-boarding process?")

```

### Q27. How could on-boarding at your company be improved?

```{r, fig.align='center', out.width = '100%', fig.height=3}

compute_perc(Qs, improve_onboarding_27 )

```

### Q28. How much is your monthly salary in Ghana Cedis?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, monthly_salary_28 ), 
              a_factor = monthly_salary_28 , its_values = perc, labels = perc_label, 
              title = "Q28. How much is your monthly salary in Ghana Cedis?")

```

### Q29. Which of the following best describes your current job-seeking status?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, job_status_29 ), 
              a_factor = job_status_29 , its_values = perc, labels = perc_label, 
              title = "Q29. Which of the following best describes your current job-seeking status?")

```

### Q30. In general, what drives you to look for a new job? Select all that apply.

#### All Respondents

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```

#### Professional Developers
```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a developer by profession"), drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```

#### Students

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs %>% filter(profession_1 == "I am a student who is learning to code"), drivers_for_new_jobs_30), 
              a_factor = drivers_for_new_jobs_30, its_values = perc, labels = perc_label, 
              title = "")

```



### Q31. When job searching, how do you learn more about a company? Select all that apply.

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, how_learn_about_company_31 ), 
              a_factor = how_learn_about_company_31 , its_values = perc, labels = perc_label, 
              title = "Q31. When job searching, how do you learn more about a company?")

```

### Q32. Are you a member of any other online/offline developer communities?

```{r, fig.align='center', out.width = '100%', fig.height=3}

basic_barplot(my_df = compute_perc(Qs, dev_community_member_32 ), 
              a_factor = dev_community_member_32 , its_values = perc, labels = perc_label, 
              title = "Q32. Are you a member of any other online/offline developer communities?")

```

### Q33. Developer communities you are a member of

```{r, fig.align='center', out.width = '100%', fig.height=3}

compute_perc(Qs, which_dev_community_33 )


```


## Skills

-> some cleaning needed.  

* duplicate names -> clean
* not same names as original -> revert to original name (no _, capitalized)
* group as Hacklab did 
* 

```{r}
skills %>% count(tool)
```


### Q.8 Which programming, scripting and markup languages have you done extensive development work in over the past year, and which do you want to work in over the next year? (If you both worked with the language and want to continue to do so, please check both)


```{r, fig.align='center', out.width = '100%', fig.height=6}

languages = c(
  "C",			
  "C#",			
  "C++",		
  "HTML/CSS",			
  "Java",			
  "JavaScript",			
  'Kotlin',			
  "PHP",			
  "Python",			
  "Ruby",			
  "SQL",			
  "Swift",			
  "TypeScript"
)

skills  %>% filter(tool %in% languages) %>% 
  group_by(tool, level) %>%
  summarize(n = n()) %>% 
  arrange(tool, level)


```














