---
title: "TITEL"
output:
  html_document:
    theme: darkly # select theme, I like darkly
    toc: true     # adding table of contents
    toc_float: true # floating table of contents
---
```{css, echo=FALSE}
fluid-row .title{
  font-size: 10vm;
  color: #FFFFFF;
}

p {                 
  font-size: 1vw;
  font-family: "Gotham", Gotham, sans-serif;
}

body {               
  background-color: #610b70;
}

h1 {             
  font-size: 5vw;
  font-weight: bold;
  color: #88b101;
}

h2 {
  font-size: 3vw;
  color: #eb1c96;
}

h3 {
  font-size: 3vw;
  color: #FFFFFF;
}

```
<!-- 
highest header (h1) styling (Header indicated by #)
subheader (h2) styling (Header indicated by ##)
body = background
vw is for responsive font size
-->


```{table of contents, echo=FALSE}
#toc_container {
    background: #f9f9f9 none repeat scroll 0 0;
    border: 1px solid #aaa;
    display: table;
    font-size: 100px;
    margin-bottom: 1em;
    padding: 20px;
    width: auto;
}
.toc_title {
    font-weight: 700;
    text-align: center;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bslib)
```
# Hacklab Report

## Themes
Themes to choosefrom: <br>
flatly, darkly, readable, spacelab, united, cosmo, lumen, paper, sandstone, simplex, and yeti.

```{r cars}

```

### Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Another header 1 that is really really loooooong
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>

## Another header 1
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
## Another header 1
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>

## Another header 1
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>

## Another header 1
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>
Text text <br>

### Data Overview {.tabset .tabset-fade .tabset-pills .pad}
```{css tab colors, echo=FALSE}
/* Style the tab */
.tab {
  float: left;
  border: 1px solid #ccc;
  background-color: #f1f1f1;
  width: 30%;
  height: 300px;
}

/* Style the buttons that are used to open the tab content */
.tab button {
  display: block;
  background-color: inherit;
  color: black;
  padding: 22px 16px;
  width: 100%;
  border: none;
  outline: none;
  text-align: left;
  cursor: pointer;
}

/* Change background color of buttons on hover */
.tab button:hover {
  background-color: #ddd;
}

/* Create an active/current "tab button" class */
.tab button.active {
  background-color: #ccc;
}

/* Style the tab content */
.tabcontent {
  float: left;
  padding: 0px 12px;
  border: 1px solid #ccc;
  width: 70%;
  border-left: none;
  height: 300px;
  display: none;
}
```

<!-- tabset-pills is the background color of the button
.tabset-fade indicates whether the button should fade when clicking on the button
.tabset is the command to create a button/tab
-->

Glimpse and Summary give us a quick overview. Helpers are functions we'll use later on. 

<br><br>

#### Button1

Although the number of unique job ads is fairly modest, there is some very rich data to explore. We have over a year's worth of data, ranging between

```{r}

```


#### Button2

Summary for each variable (only included R and Python here, as other languages (along with R and Python) are explored more deeply later.

```{r}

```


#### Button3

Some functions are defined here for later use.

#### Button4 with subtabs {.tabset}
##### Tab Nr 1{.tabset}
(tab1 content)

###### Subtab
(subtab content)

##### Tab Nr 2
(tab2 content)
<!-- Important to note: Buttons need to be one level below the title
-->

